### 2016-05-10 paper/paper.tex
Pushing latex

\title{Probabilistic Audio Localization in Wireless Acoustic Sensor Networks}

\date{\today}

\author{Alexander Wallar}

\documentclass[12pt]{article}

\usepackage[pdftex]{graphicx}

\usepackage{amsmath}

\usepackage{relsize}

\usepackage{float}

\usepackage{algorithm}

\usepackage[noend]{algorithmic}

\floatstyle{ruled} \newfloat{program}{thp}{lop} \floatname{program}{Structure}

\newcommand{\Normal}[3]{\mathcal{N}(#1, #2, #3)}

\newcommand{\Acronym}[1]{\ensuremath{{\small{\texttt{#1}}}}}
\newcommand{\Name}{\Acronym{Camgaze.js}} \newcommand{\False}{\Constant{false}}
\newcommand{\True}{\Constant{true}}
\newcommand{\Symbol}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\Function}[1]{\ensuremath{{\small \textsc{#1}}}}
\newcommand{\Constant}[1]{\ensuremath{\small{\texttt{#1}}}}
\newcommand{\Var}[1]{\ensuremath{{\small{\textsl{#1}}}}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
\newcommand{\grad}[1]{\underset{#1}{\operatorname{\Function{GradientDecent}}}\;}

\begin{document}

\maketitle

\begin{algorithm}[ht] 
    \caption{Multiple Source Sound Localization}
    \label{algo:MS}
    \begin{algorithmic}[1]
        \setcounter{ALC@line}{0}
        \vspace*{1mm}

        \STATE $\Function{GetLocations}(\Var{r}, \Var{spl}, \Var{eventList}) \rightarrow$
        \STATE $\Var{locations} \leftarrow \Function{InitLocationList}()$
        \FORALL {$e \in \Var{eventList}$}
        \STATE $\Var{locations}.\Var{add}(\mathcal{L}_m(\Var{r}, \Var{spl}, \Var{eventList}, e.\Var{x}, e.\Var{y}))$
        \ENDFOR
        \STATE $\Var{centers} \leftarrow \Function{AffinityPropagation}(\Var{locations})$
        \COMMENT {\Var{centers} is a list of $(x, y, c)$ points where $c$ is the confidence}
        \RETURN $\Var{centers}$

    \end{algorithmic}
\end{algorithm}

\bibliographystyle{ieeetran}
\bibliography{bib}

\end{document}

### 2016-05-10 paper/paper.tex

@@
 
+Dimensionality reduction is a key component of making robotic motion planning
+fast and efficient. A rigid body has six degrees of freedom (DoF), so a
+multi-link robot such as a humanoid or snake robot can have dozens or hundreds
+of DoF. The rotational and translational transformations of a rigid body can be
+described with the 3-dimensional Special Euclidean group (known as SE(3)).
+SE(3) is homeomorphic to the topological space R3xRP3 (where RP3 is the
+3-dimensional real projective plane). Therefore it is easy to imagine a
+multi-link robot with an arbitrarily complex state space. The picture is
+further complicated by the addition of obstacles into a robot's world. The
+state space for motion planning, which must take into account both the robot
+and obstacles, is called the configuration space.
+
+Motion planning essentially consists of a search over the configuration space
+from a start configuration to a goal configuration. A huge number of methods
+for searching through the configuration space have been developed, most of
+which can be divided into two classes, sampling-based motion planning and
+combinatorial motion planning. In sampling-based motion planning, obstacles in
+the C-space are defined implicitly so that the path is constructed by randomly
+or pseudorandomly sampling points from the C-space and doing collision
+checking. In combinatorial motion planning, obstacles are defined explicitly
+and a complete search of the C-space is made. Because of the difficulty of
+explicitly defining all of the geometry in a potentially complex world,
+sampling-based motion planning techniques such as the Rapidly-exploring Random
+Tree (RRT) and Probabilistic RoadMap (PRM) planning are the most popular motion
+planning techniques.
+
+[RRT* algorithm]
+[PRM* algorithm]
+
+We used RRT* and PRM* in our project. The * indicates that these are the
+asymptotically optimal versions of RRT and PRM (they are guaranteed to find the
+optimal path as time goes to infinity). [Description of how RRT* and PRM* work]
+
+Despite the efficiency of sampling-based search, the high dimensionality of the
+configuration space of complex robots means that motion planning for many
+robots is slow or infeasible. This is because of the "curse of dimensionality",
+in which the size of the search space increases exponentially with the
+dimension. One way to speed up motion planning is to reduce the dimensionality
+of the C-space and do search in this reduced dimensionality space. For this
+project, we tested the performance of a number of dimensionality reduction
+techniques to see which one was the most effective in speeding up motion
+planning while providing reliable obstacle avoidance.
+
+We used FeatureAgglomeration, TruncatedSVD, PCA, KernelPCA, RandomizedPCA, and
+our own "Trained Johnson-Lindestrauss" to perform dimensionality reduction.
+
+PCA: Principle Component Analysis, or PCA, transforms a set of variables in
+their "principle components", or linearly uncorrelated variables.
+
+ncollisions, duration, length, num failed. Run each one 100 times
+
+\cite{wallar2015reactive}
+
 \begin{algorithm}[ht] 
@@
 
-\bibliographystyle{ieeetran}
+\bibliographystyle{plain}
 \bibliography{bib}

### 2016-05-10 paper/paper.tex
PRM Star in latex done
paper changes

@@
 
+The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations.The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the nearest existing vertices in the roadmap.
+in
+
 Despite the efficiency of sampling-based search, the high dimensionality of the
@@
 \begin{algorithm}[ht] 
-    \caption{Multiple Source Sound Localization}
-    \label{algo:MS}
+    \caption{PRM*}
+    \label{algo:PRMstar}
     \begin{algorithmic}[1]

### 2016-05-10 paper/paper.tex
Pushing data

--- 2016-05-10 paper/paper.tex
@@
-\title{Probabilistic Audio Localization in Wireless Acoustic Sensor Networks}
+\title{A Comparison of Dimensionality Reduction Techniques for Robotic Motion
+Planning}
 
@@
 
-\author{Alexander Wallar}
+\author{Brandon Araki \and Alex Wallar}
 
@@
 \newcommand{\grad}[1]{\underset{#1}{\operatorname{\Function{GradientDecent}}}\;}
+
+\newcommand{\xr}[0]{\ensuremath{{x_{\Var{rand}}}}}
+\newcommand{\xnear}[0]{\ensuremath{{x_{\Var{near}}}}}
+\newcommand{\xnearest}[0]{\ensuremath{{x_{\Var{nearest}}}}}
+\newcommand{\xnew}[0]{\ensuremath{{x_{\Var{new}}}}}
+\newcommand{\xmin}[0]{\ensuremath{{x_{\Var{min}}}}}
+\newcommand{\cmin}[0]{\ensuremath{{c_{\Var{min}}}}}
+\newcommand{\cost}[0]{\ensuremath{{\Function{Cost}}}}
+\newcommand{\linef}[1]{\ensuremath{{c(\Function{Line}(#1))}}}
 
@@
 PCA: Principle Component Analysis, or PCA, transforms a set of variables in
-their "principle components", or linearly uncorrelated variables.
+their "principle components", or linearly uncorrelated variables.\
 
@@
 
-\cite{wallar2015reactive}
-
 \begin{algorithm}[ht] 
-    \caption{Multiple Source Sound Localization}
+    \caption{$\Function{PRM}^*$~\cite{karaman2011sampling}}
     \label{algo:MS}
@@
 
-        \STATE $\Function{GetLocations}(\Var{r}, \Var{spl}, \Var{eventList}) \rightarrow$
-        \STATE $\Var{locations} \leftarrow \Function{InitLocationList}()$
-        \FORALL {$e \in \Var{eventList}$}
-        \STATE $\Var{locations}.\Var{add}(\mathcal{L}_m(\Var{r}, \Var{spl}, \Var{eventList}, e.\Var{x}, e.\Var{y}))$
+        \STATE $V \leftarrow \{x_{\Var{init}}\} \cup \{\Var{SampleFree}_i\}_{i=1,\ldots,n}$
+        \STATE $E \leftarrow \emptyset$
+        \FORALL{$v \in V$}
+            \STATE $U \leftarrow \Function{Near}(G = (V, E), v,
+            \gamma \cdot \Function{PRM}(\log(n) / n) ^ {1 / d})
+            \backslash \{v\}$
+            \FORALL{$u \in U$}
+                \IF{$\Function{CollisionFree}(v, u)$}
+                    \STATE $E \leftarrow E \cup \{(v, u), (u, v)\}$
+                \ENDIF
             \ENDFOR
-        \STATE $\Var{centers} \leftarrow \Function{AffinityPropagation}(\Var{locations})$
-        \COMMENT {\Var{centers} is a list of $(x, y, c)$ points where $c$ is the confidence}
-        \RETURN $\Var{centers}$
+        \ENDFOR
+        \RETURN $G = (V, E)$
+
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[ht] 
+    \caption{$\Function{RRT}^*$~\cite{karaman2011sampling}}
+    \label{algo:MS}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+
+        \STATE $V \leftarrow \{x_{\Var{init}}\}$
+        \STATE $E \leftarrow \emptyset$
+        \FOR{$i = 1 \ \TO \ n$}
+            \STATE $\xr \leftarrow \Var{SampleFree}_i$
+            \STATE $\xnearest \leftarrow \Function{Nearest}(G = (V, E), \xr)$
+            \STATE $\xnew \leftarrow \Function{Steer}(\xnearest, \xr)$
+            \IF{$\Function{ObstacleFree}(\xnearest, \xnew)$}
+                \STATE $\xnear \leftarrow \Function{Near}(G = (V, E), \xnew,
+                \min \{\gamma \cdot \Function{RRT}(\log(|V|) / |V|)^{1 / d}, \eta\})$
+                \STATE $V \leftarrow V \cup \{\xnew\}$
+                \STATE $\xmin \leftarrow \xnearest$
+                \STATE $\cmin \leftarrow \Function{Cost}(\xnear) +
+                c(\Function{Line}(\xnearest, \xnew))$
+                \FORALL{$\xnear \in X_{\Var{near}}$}
+                    \STATE $C \leftarrow \cost(\xnear) + \linef{\xnear, \xnew}$
+                    \IF{$\Function{CollisionFree}(\xnear, \xnew) \wedge C < \cmin$}
+                        \STATE $\xmin \leftarrow \xnear$
+                        \STATE $\cmin \leftarrow C$
+                    \ENDIF
+                    \STATE $E \leftarrow E \cup \{(\xmin, \xnew)\}$
+                \ENDFOR
+            \ENDIF
+        \ENDFOR
+        \RETURN $G = (V, E)$
 

### 2016-05-10 scripts/plot_metrics.py

@@ -3,6 +3,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import numpy as np
 import warnings
+import pandas
 from collections import defaultdict
 
 
@@ -19,10 +20,12 @@ field_names = ["planner", "transformer", "n_collisions",
 
 def load_data(filename):
     data = np.genfromtxt(filename, delimiter=",", dtype=None)[1:]
-    planner_data = recursive_defaultdict(3, float)
+    planner_data = recursive_defaultdict(4, float)
     for row in data:
-        for i in xrange(len(field_names) - 2):
-            planner_data[row[0]][row[1]][field_names[i + 2]] = row[i + 2]
+        for i in xrange(len(field_names) - 3):
+            planner_data[row[0]][
+                row[1]][row[-1]][field_names[i + 2]] = row[i + 2]
+    print planner_data
     return planner_data
 
 
@@ -38,21 +41,22 @@ def recursive_defaultdict_helper(dd, level):
         return recursive_defaultdict_helper(new_dd, level - 1)
 
 
-def get_all(data, pl, key):
+def get_all(data, pl, n_ob, key):
     xs = list()
     for tr in transformer_strs:
-        xs.append(data[pl][tr][key])
+        xs.append(data[pl][tr][n_ob][key])
+    print xs
     return xs
 
 
-def make_plots(data, key):
+def make_plots(data, key, n_ob):
     rects = list()
     fig, ax = plt.subplots()
     ind = np.arange(len(transformer_strs))
     width = 0.35
     for i, pl in enumerate(planner_strs):
         rect = ax.bar(
-            ind + i * width, get_all(data, pl, key), width,
+            ind + i * width, get_all(data, pl, n_ob, key), width,
             color=pl_clrs[pl],
             label=pl)
         rects.append(rect)
@@ -64,6 +68,8 @@ def make_plots(data, key):
 if __name__ == "__main__":
     warnings.simplefilter("ignore")
     sns.set_context("poster", font_scale=2.2)
-    data = load_data("data/data.csv")
-    make_plots(data, "n_collisions")
+    data = pandas.read_csv("data/data.csv")
+    # subdata = data[data["n_obs"] == 30]
+    sns.barplot(x="transformer", y="path_length", hue="n_obs", data=data)
+    # make_plots(data, "n_collisions", 20)
     plt.show()

### 2016-05-10 tests/experiments.py

@@ -34,6 +34,7 @@ n_runs = 100
 high_dim = 6
 low_dim = 4
 timeout = 1.5
+rad_mean = 7
 
 
 def run_experiments(filename):

### 2016-05-10 tests/test_planner.py

@@ -11,7 +11,7 @@ from ompl import geometric as og
 
 
 def sanity_check():
-    obs = flatland.RandomObstacleGen(dim=3).generate(30)
+    obs = flatland.RandomObstacleGen(dim=3, rad_mean=7).generate(30)
     start = np.array([-10, -10, -10])
     goal = np.array([10, 10, 10])
     planner = flatland.FLPlanner(

### 2016-05-10 paper/paper.tex

@@
 optimal path as time goes to infinity). [Description of how RRT* and PRM* work]
+
+The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations.The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the nearest existing vertices in the roadmap.
+in
 

### 2016-05-10 scripts/plot_metrics.py
RRT star done

@@ -68,7 +68,7 @@ def make_plots(data, key, n_ob):
 if __name__ == "__main__":
     warnings.simplefilter("ignore")
     sns.set_context("poster", font_scale=2.2)
-    data = pandas.read_csv("data/data.csv")
+    data = pandas.read_csv("data/data_copy.csv")
     # subdata = data[data["n_obs"] == 30]
     sns.barplot(x="transformer", y="path_length", hue="n_obs", data=data)
     # make_plots(data, "n_collisions", 20)

### 2016-05-10 tests/experiments.py

@@ -29,12 +29,12 @@ field_names = ["planner", "transformer", "n_collisions",
                "n_collisions_std", "planning_duration_std", "path_length_std",
                "num_failed", "n_obs"]
 
-n_obs = [10, 20, 30]
+n_obs = [30]
 n_runs = 100
 high_dim = 6
 low_dim = 4
 timeout = 1.5
-rad_mean = 7
+rad_mean = 6.5
 
 
 def run_experiments(filename):

### 2016-05-11 paper/paper.tex
RRT star done
added to paper

--- 2016-05-10 paper/paper.tex
@@
 state space for motion planning, which must take into account both the robot
-and obstacles, is called the configuration space.
+and obstacles, is called the configuration space \cite{lavalle2006planning}.
 
@@
 
-[RRT* algorithm]
-[PRM* algorithm]
-
 We used RRT* and PRM* in our project. The * indicates that these are the
 asymptotically optimal versions of RRT and PRM (they are guaranteed to find the
-optimal path as time goes to infinity). [Description of how RRT* and PRM* work]
+optimal path as time goes to infinity).
 
-The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations.The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the nearest existing vertices in the roadmap.
-in
+The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations.The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
+PRM can be elevated to the asymptotically optimal PRM* by defining the radius to be a function of the number of vertices and the dimensionality of the space.
+
+The Rapidly-expanding Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration.
+Starting at the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
+RRT* differs from RRT in that whenever a new node is added to the tree, connections are also tested to the new node from every other node that is within a certain radius from the new node.
+The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.
 
@@
 
-\cite{wallar2015reactive}
-
 \begin{algorithm}[ht] 
-    \caption{PRM*}
-    \label{algo:PRMstar}
+    \caption{$\Function{PRM}^*$~\cite{karaman2011sampling}}
+    \label{algo:MS}
     \begin{algorithmic}[1]
@@
 
-        \STATE $\Function{GetLocations}(\Var{r}, \Var{spl}, \Var{eventList}) \rightarrow$
-        \STATE $\Var{locations} \leftarrow \Function{InitLocationList}()$
-        \FORALL {$e \in \Var{eventList}$}
-        \STATE $\Var{locations}.\Var{add}(\mathcal{L}_m(\Var{r}, \Var{spl}, \Var{eventList}, e.\Var{x}, e.\Var{y}))$
+        \STATE $V \leftarrow \{x_{init}\} \cup \{\Var{SampleFree}_i\}_{i=1,\ldots,n}$
+        \STATE $E \leftarrow \emptyset$
+        \FORALL{$v \in V$}
+            \STATE $U \leftarrow \Function{Near}(G = (V, E), v,
+            \gamma \cdot \Function{PRM}(\log(n) / n) ^ {1 / d})
+            \backslash \{v\}$
+            \FORALL{$u \in U$}
+                \IF{$\Function{CollisionFree}(v, u)$}
+                    \STATE $E \leftarrow E \cup \{(v, u), (u, v)\}$
+                \ENDIF
             \ENDFOR
-        \STATE $\Var{centers} \leftarrow \Function{AffinityPropagation}(\Var{locations})$
-        \COMMENT {\Var{centers} is a list of $(x, y, c)$ points where $c$ is the confidence}
-        \RETURN $\Var{centers}$
+        \ENDFOR
+        \RETURN $G = (V, E)$
 

### 2016-05-11 paper/paper.tex

--- 2016-05-10 paper/paper.tex
@@
 \newcommand{\xmin}[0]{\ensuremath{{x_{\Var{min}}}}}
+\newcommand{\xparent}[0]{\ensuremath{{x_{\Var{parent}}}}}
 \newcommand{\cmin}[0]{\ensuremath{{c_{\Var{min}}}}}
@@
 
-[RRT* algorithm]
-[PRM* algorithm]
-
 We used RRT* and PRM* in our project. The * indicates that these are the
 asymptotically optimal versions of RRT and PRM (they are guaranteed to find the
-optimal path as time goes to infinity). [Description of how RRT* and PRM* work]
+optimal path as time goes to infinity).
+
+The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations.The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
+PRM can be elevated to the asymptotically optimal PRM* by defining the radius to be a function of the number of vertices and the dimensionality of the space.
+
+The Rapidly-expanding Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration.
+Starting at the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
+RRT* differs from RRT in that whenever a new node is added to the tree, connections are also tested to the new node from every other node that is within a certain radius from the new node.
+The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.
 
@@
                 \ENDFOR
+                \FORALL{$\xnear \in X_{\Var{near}}$}
+                    \STATE $C \leftarrow \cost(\xnew) + \linef{\xnew, \xnear}$
+                    \IF{$\Function{CollisionFree}(\xnew, \xnear) \wedge C < \cost(\xnear)$}
+                        \STATE $\xparent \leftarrow \Function{Parent}(\xnear)$
+                        \STATE $E \leftarrow (E \backslash \{(\xparent, \xnear)\}) \cup \{(\xnew, \xnear)\}$
+                    \ENDIF
+                \ENDFOR
             \ENDIF

### 2016-05-11 paper/paper.tex
added to paper

@@
 
-We used FeatureAgglomeration, TruncatedSVD, PCA, KernelPCA, RandomizedPCA, and
-our own "Trained Johnson-Lindestrauss" to perform dimensionality reduction.
+We used Feature Agglomeration, Truncated SVD, PCA, Kernel PCA, Randomized PCA, and
+our own "Trained Johnson-Lindestrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
 
-PCA: Principle Component Analysis, or PCA, transforms a set of variables in
-their "principle components", or linearly uncorrelated variables.\
+PCA: Principle Component Analysis, or PCA, transforms a set of variables into 
+their "principle components", or linearly uncorrelated variables.
+In other words, the axes of the dataset in question are reoriented so the Rayleigh quotient, which is related to the covariance of the dataset, along the first axis is maximized.
+After the first axis has been set, the second axis is then set to maximize the Rayleigh quotient of the dataset along that axis, and so on, so that the first axis has the highest Rayleigh quotient, the second axis has the second highest, and so on.
+Dimensionality reduction is then performed by chopping off the axes with the lowest Rayleigh quotients (which corresponded to axes with the lowest variance).
+\
+
+Kernel PCA: Kernel PCA enables non-linear dimensionality reduction by using kernels to create a non-linear mapping from a high dimension to a lower dimension.
+
+Randomized PCA: In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
+
+Feature Agglomeration: Feature Agglomeration groups together similar features to reduce the number of features.
+It does this by using agglomerative clustering, in which every data point starts as a cluster, and clusters are merged together.
+
+Truncated SVD:
 

### 2016-05-11 paper/paper.tex
added to paper

@@
 We used Feature Agglomeration, Truncated SVD, PCA, Kernel PCA, Randomized PCA, and
-our own "Trained Johnson-Lindestrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
+our own "Trained Johnson-Lindenstrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
 
@@
 
-Truncated SVD:
+Truncated SVD: Truncated SVD works best with sparse matrices.
+
+Trained Johnson-Lindenstrauss: Normal Johnson-Lindenstrauss uses a random mxn Gaussian matrix to reduce the dimensionality of a vector from n to m.
+Instead of using a random matrix, we train a matrix on randomly-generated obstacles in n-space to minimize [MINIMIZE WHAT?]
+
+Methods:
 

### 2016-05-11 flatland/obstacles.py
balls

@@ -15,8 +15,8 @@ class RandomObstacleGen(object):
         self.rad_mean = kwargs.get("rad_mean", 5)
         self.rad_std = kwargs.get("rad_std", 0.5)
         self.dim = kwargs.get("dim", 2)
-        self.sample_pts_mean = kwargs.get("sample_pts_mean", self.dim + 10)
-        self.sample_pts_std = kwargs.get("sample_pts_std", 3)
+        self.sample_pts_mean = kwargs.get("sample_pts_mean", self.dim ** 2)
+        self.sample_pts_std = kwargs.get("sample_pts_std", 0.1)
 
     def sample_n_sphere(self, rad, center):
         pt = np.zeros((self.dim,))
@@ -50,6 +50,6 @@ class RandomObstacleGen(object):
             ch = spatial.ConvexHull(pts)
             eqs = ch.equations
             poly = polytope.Polytope(
-                eqs[:, : - 1], -eqs[:, -1])
+                eqs[:, : - 1], -eqs[:, -1], minrep=True)
             obs.append((poly, ch.points))
         return obs

### 2016-05-11 flatland/planner.py

@@ -66,8 +66,8 @@ class FLPlanner(object):
     def solve(self, st, gl, timeout=1.0):
         space = ob.RealVectorStateSpace(self.dim)
         bounds = ob.RealVectorBounds(self.dim)
-        bounds.setLow(self.low_bound)
-        bounds.setHigh(self.high_bound)
+        bounds.setLow(min(st.min(), gl.min()))
+        bounds.setHigh(max(st.max(), gl.max()))
         space.setBounds(bounds)
         si = ob.SpaceInformation(space)
         si.setStateValidityChecker(

### 2016-05-11 flatland/plot_path.py

@@ -60,6 +60,7 @@ def make_path_plot_2d():
 
 def make_path_plot_3d():
     fig1 = plt.figure()
+    ax = fig1.gca(projection='3d')
 
     for i in os.listdir(obstacleFolder):
         obstaclePath = obstacleFolder + i
@@ -69,7 +70,6 @@ def make_path_plot_3d():
         CH = Delaunay(ob).convex_hull
         x,y,z = ob[:,0],ob[:,1],ob[:,2]
 
-        ax = fig1.gca(projection='3d')
         S = ax.plot_trisurf(x,y,z,triangles=CH,shade=True,cmap=cm.copper,lw=0.2)
 
     txt = np.loadtxt(rrtPath)

### 2016-05-11 paper/paper.tex

@@
 Dimensionality reduction is then performed by chopping off the axes with the lowest Rayleigh quotients (which corresponded to axes with the lowest variance).
-\
+
+$$
+\mathbf{w}_{(1)} = {\operatorname{\arg\,max}}\, \left\{ \frac{\mathbf{w}^T\mathbf{X}^T \mathbf{X w}}{\mathbf{w}^T \mathbf{w}} \right\}
+$$
 
@@
 
+Simplifications: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
+Therefore we limited ourselves to path planning for a point in $R^{n}$, and we approximate obstacles as being convex polytopes.
+This is a valid simplification of the robot motion planning problem for several reasons.
+The first reason is that most real robots actually operate in a C-space similar to $R^{n}$.
+This is because $RP^{3}$, the space in which the rotation of a rigid body occurs, simplifies to $R^{3}$ if the joint cannot make a full rotation.
+Second, in defining the C-space, the obstacle space takes into account the interaction of the robot geometry with obstacles.
+Therefore motion planning through the free space can be done using a point to represent the robot (since the geometry of the robot is implicitly included in the obstacle space).
+Therefore, although our simulations do not represent real robots and real obstacles, they do approximate them.
+
+We implemented an n-dimensional world in which to test these algorithms.
+We structured our code so that we could run our tests in any space of dimension d >= 2.
+We also made visualization scripts for the 2- and 3-dimensional tests.
+
+Next we made a random obstacle generator. We create n-dimensional convex polytopes by sampling $n^{2}$ random points on the n-sphere using the equations below.
+
+\begin{align}
+x_1 &= r \cos(\phi_1) \\
+x_2 &= r \sin(\phi_1) \cos(\phi_2) \\
+x_3 &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \\
+    &\vdots\\
+x_{n-1} &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \cos(\phi_{n-1}) \\
+x_n &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
+\end{align}
+
+We then run Feature Agglomeration, Truncated SVD, PCA, Kernel PCA, and Randomized PCA to reduce the dimensionality of the space.
+We then run PRM* on each of these reduced dimensionality spaces, as well as on the full-dimensionality space as a control.
+We then run an inverse transform on the paths that we find to return them to the full-dimensionality space.
+Then
+
 ncollisions, duration, length, num failed. Run each one 100 times
+
+\begin{algorithm}[ht] 
+    \caption{$\Function{Dimensionality Reduction}$}
+    \label{algo:MS}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+
+        \STATE $Specify dimension, dimensionality reduction technique, and obstacle information$
+        \STATE $Generate obstacles based off of obstacle information$
+        \STATE $Fit dimensionality reduction transform to the space$
+        \STATE $Transform the space$
+        \STATE $Run PRM* over the space$
+        \STATE $Inverse transform the path back to the original dimension$
+        \STATE $Check for collisions$
+    \end{algorithmic}
+\end{algorithm}
 

### 2016-05-11 scripts/plot_path.py
removed old plots

@@ -1,41 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.path import Path
-import matplotlib.patches as patches
-import os
-import pdb
-
-
-rrtPath = "sandbox/prmpath.txt"
-obstacleFolder = 'sandbox/obstacles/'
-
-
-def generate_codes(length):
-	codes = list()
-	codes.append(Path.MOVETO)
-	for _ in xrange(1, length - 1):
-		codes.append(Path.LINETO)
-	codes.append(Path.STOP)
-	return codes
-
-
-def make_path_plot():
-	fig1 = plt.figure()
-
-	for i in os.listdir(obstacleFolder):
-	    obstaclePath = obstacleFolder + i
-
-	    ob = np.loadtxt(obstaclePath)
-	    ob = np.vstack((ob, ob[0, :]))
-	    path = Path(ob)
-	    patch = patches.PathPatch(path, facecolor='orange', lw=2)
-	    ax = plt.gca()
-	    ax.add_patch(patch)
-
-	txt = np.loadtxt(rrtPath)
-	plt.plot(txt[:, 0], txt[:, 1], 'b-')
-
-
-if __name__ == "__main__":
-	make_path_plot()
-	plt.show()

### 2016-05-11 flatland/obstacles.py
Adding obstacles

@@ -15,7 +15,7 @@ class RandomObstacleGen(object):
         self.rad_mean = kwargs.get("rad_mean", 5)
         self.rad_std = kwargs.get("rad_std", 0.5)
         self.dim = kwargs.get("dim", 2)
-        self.sample_pts_mean = kwargs.get("sample_pts_mean", self.dim ** 2)
+        self.sample_pts_mean = kwargs.get("sample_pts_mean", self.dim + 10)
         self.sample_pts_std = kwargs.get("sample_pts_std", 0.1)
 
     def sample_n_sphere(self, rad, center):

### 2016-05-11 flatland/plot_path.py

@@ -33,8 +33,9 @@ def make_path_plot_2d():
 
         ob = np.loadtxt(obstaclePath)
         ob = np.vstack((ob, ob[0, :]))
-        #CH = Delaunay(ob).convex_hull
-        #x,y,z = ob[:,0],ob[:,1]
+        CH = Delaunay(ob).convex_hull
+        x,y = ob[:,0],ob[:,1]
+        z = np.zeros_like(x)
 
         ax = fig1.gca(projection='3d')
         S = ax.plot_trisurf(x,y,z,triangles=CH,shade=True,cmap=cm.copper,lw=0.2)

### 2016-05-11 flatland/plot_path.py
stuff

@@ -36,8 +36,8 @@ def make_path_plot_2d():
         #CH = Delaunay(ob).convex_hull
         #x,y,z = ob[:,0],ob[:,1]
 
-        ax = fig1.gca(projection='3d')
-        S = ax.plot_trisurf(x,y,z,triangles=CH,shade=True,cmap=cm.copper,lw=0.2)
+        #ax = fig1.gca(projection='3d')
+        #S = ax.plot_trisurf(x,y,z,triangles=CH,shade=True,cmap=cm.copper,lw=0.2)
 
         path = Path(ob)
         patch = patches.PathPatch(path, facecolor='orange', lw=2)

### 2016-05-11 flatland/plot_path.py
fixing brandons shit

@@ -91,5 +91,5 @@ def make_path_plot_3d():
     plt.plot(txt[:, 0], txt[:, 1], txt[:,2], 'bo-')
 
 if __name__ == "__main__":
-    make_path_plot_2d()
+    make_path_plot_3d()
     plt.show()

### 2016-05-11 tests/experiments.py

@@ -31,12 +31,13 @@ field_names = ["planner", "transformer", "n_collisions",
                "num_failed", "n_obs", "is_full_dim"]
 
 # n_obs = [70, 90, 110]
-n_obs = [3]
+n_obs = [14]
 n_runs = 2
-high_dim = 2
-low_dim = 2
+high_dim = 3
+low_dim = 3
 timeout = 1.5
-rad_mean = 7
+rad_mean = 5
+# rad_mean = 7
 
 
 def path_size(path):

### 2016-05-11 flatland/plot_path.py
adding 3d example figs

@@ -81,15 +81,19 @@ def make_path_plot_3d():
     start = txt[0,0:2]
     goal = txt[-1,0:2]
 
-    startCircle = Circle(start,0.3,color='g')
-    goalCircle = Circle(goal,0.3,color='r')
+    startCircle = Circle(start,2,color='g')
+    goalCircle = Circle(goal,2,color='r')
     ax.add_patch(startCircle)
     ax.add_patch(goalCircle)
-    art3d.pathpatch_2d_to_3d(startCircle,z=txt[0,2],zdir='z')
-    art3d.pathpatch_2d_to_3d(goalCircle,z=txt[-1,2],zdir='z')
+    art3d.pathpatch_2d_to_3d(startCircle,z=txt[0,2],zdir='y')
+    art3d.pathpatch_2d_to_3d(goalCircle,z=txt[-1,2],zdir='y')
 
-    plt.plot(txt[:, 0], txt[:, 1], txt[:,2], 'bo-')
+    plt.plot(txt[:, 0], txt[:, 1], txt[:,2], 'bo-', markersize=10, linewidth=3)
 
 if __name__ == "__main__":
-    make_path_plot_3d()
+    import sys
+    if sys.argv[1] == "2":
+        make_path_plot_2d()
+    else:
+        make_path_plot_3d()
     plt.show()

### 2016-05-11 tests/experiments.py

@@ -61,6 +61,7 @@ def run_experiments(filename):
         for i, tr in enumerate(transformers):
             for j, pl in enumerate(planners):
                 random.seed(0)
+                np.random.seed(0)
                 for n_ob in n_obs:
                     row = dict()
                     data = defaultdict(list)

### 2016-05-11 flatland/plot_path.py
got 2d plots to work again

@@ -7,6 +7,7 @@ import mpl_toolkits.mplot3d.art3d as art3d
 import matplotlib.patches as patches
 from matplotlib.patches import Circle
 from scipy.spatial import Delaunay
+from scipy.spatial import ConvexHull
 import matplotlib.cm as cm
 import os
 import pdb
@@ -31,24 +32,31 @@ def make_path_plot_2d():
     for i in os.listdir(obstacleFolder):
         obstaclePath = obstacleFolder + i
 
-        ob = np.loadtxt(obstaclePath)
-        ob = np.vstack((ob, ob[0, :]))
-        CH = Delaunay(ob).convex_hull
-        x,y = ob[:,0],ob[:,1]
-        z = np.zeros_like(x)
+        #ob = np.loadtxt(obstaclePath)
+        #ob = np.vstack((ob, ob[0, :]))
+        #CH = Delaunay(ob).convex_hull
+        #x,y = ob[:,0],ob[:,1]
+        #z = np.zeros_like(x)
 
         #ax = fig1.gca(projection='3d')
-        #S = ax.plot_trisurf(x,y,z,triangles=CH,shade=True,cmap=cm.copper,lw=0.2)
+        #S = ax.plot_trisurf(x,y,z,triangles=CH,shade=True,cmap=cm.autumn,lw=0.2)
 
-        path = Path(ob)
+        ob = np.loadtxt(obstaclePath)
+        ob = np.vstack((ob, ob[0, :]))
+        hull = ConvexHull(ob)
+        newHull = []
+        for p in hull.vertices:
+            newHull.append(ob[p])
+        newHull.append(ob[hull.vertices[0]])
+        path = Path(newHull)
         patch = patches.PathPatch(path, facecolor='orange', lw=2)
-        ax = fig1.gca(projection='3d')
+        ax = plt.gca()
         ax.add_patch(patch)
-        art3d.pathpatch_2d_to_3d(patch,z=0,zdir='z')
+
 
     txt = np.loadtxt(rrtPath)
-    start = txt[0,0:1]
-    goal = txt[-1,0:1]
+    start = txt[0,0:2]
+    goal = txt[-1,0:2]
 
     #startCircle = plt.Circle(start,0.3,color='g', zorder=2)
     #goalCircle = plt.Circle(goal,0.3,color='r', zorder=2)
[…]

### 2016-05-11 paper/paper.tex
moved figs folder from sandbox to paper
added citations

@@
 \usepackage[pdftex]{graphicx}
+\graphicspath{ {figs/} }
+\usepackage{subcaption}
 
@@
 3-dimensional real projective plane). Therefore it is easy to imagine a
-multi-link robot with an arbitrarily complex state space. The picture is
+multi-link robot with an arbitrarily complex state space \cite{transeth2008snakerobot}. The picture is
 further complicated by the addition of obstacles into a robot's world. The
 state space for motion planning, which must take into account both the robot
-and obstacles, is called the configuration space \cite{lavalle2006planning}.
+and obstacles, is called the configuration space \cite{lozano1983spatial}.
 
@@
 which can be divided into two classes, sampling-based motion planning and
-combinatorial motion planning. In sampling-based motion planning, obstacles in
+combinatorial motion planning \cite{lavalle2006planning}. In sampling-based motion planning, obstacles in
 the C-space are defined implicitly so that the path is constructed by randomly
@@
 
-The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations.The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
-PRM can be elevated to the asymptotically optimal PRM* by defining the radius to be a function of the number of vertices and the dimensionality of the space.
+The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations \cite{kavraki1996prm}.
+The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
+PRM can be elevated to the asymptotically optimal PRM* by defining the radius to be a function of the number of vertices and the dimensionality of the space \cite{karaman2011sampling}.
 
-The Rapidly-expanding Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration.
+The Rapidly-expanding Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
 Starting at the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
-RRT* differs from RRT in that whenever a new node is added to the tree, connections are also tested to the new node from every other node that is within a certain radius from the new node.
+RRT* differs from RRT in that whenever a new node is added to the tree, connections are also tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
 The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.
@@
 We then run an inverse transform on the paths that we find to return them to the full-dimensionality space.
-Then
+Then we do collision checking and check the number of collisions the path makes with the obstacles.
 
-ncollisions, duration, length, num failed. Run each one 100 times
+Results:
+
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{3d_example_1.pdf}
+        \caption{a}
+    \end{subfigure}%
+    ~ 
+    \begin{subfigure}[t]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{3d_example_2.pdf}
+        \caption{b}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{3d_example_3.pdf}
+        \caption{c}
+    \end{subfigure}
+    \caption{3d examples}
+    \label{fig:3d_examples}
+\end{figure}
+
+You can see examples of our code running in Fig ~\ref{fig:3d_examples}. The results are listed in Figure WHAT. They show that
 
@@
     \caption{$\Function{Dimensionality Reduction}$}
-    \label{algo:MS}
+    \label{algo:DR}
     \begin{algorithmic}[1]

### 2016-05-11 paper/paper.tex
added citations

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{mathabx}
 
@@
 
+\section{Background}
+
 Dimensionality reduction is a key component of making robotic motion planning
@@
 described with the 3-dimensional Special Euclidean group (known as SE(3)).
-SE(3) is homeomorphic to the topological space R3xRP3 (where RP3 is the
+SE(3) is homeomorphic to the topological space $\mathbb{R}^3\bigtimes\mathbb{RP}^3$ (where $\mathbb{RP}^3$ is the
 3-dimensional real projective plane). Therefore it is easy to imagine a
-multi-link robot with an arbitrarily complex state space \cite{transeth2008snakerobot}. The picture is
+multi-link robot \cite{kuindersma2015optimization} or a multi-robot system \cite{alonso2015multi} with an arbitrarily complex state space. The picture is
 further complicated by the addition of obstacles into a robot's world. The
@@
 
+\section{Problem Description}
+
 Despite the efficiency of sampling-based search, the high dimensionality of the
 configuration space of complex robots means that motion planning for many
-robots is slow or infeasible. This is because of the "curse of dimensionality",
+robots is slow or infeasible. This is because of the ``curse of dimensionality",
 in which the size of the search space increases exponentially with the
@@
 We used Feature Agglomeration, Truncated SVD, PCA, Kernel PCA, Randomized PCA, and
-our own "Trained Johnson-Lindenstrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
+our own ``Trained Johnson-Lindenstrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
 
 PCA: Principle Component Analysis, or PCA, transforms a set of variables into 
-their "principle components", or linearly uncorrelated variables.
+their ``principle components", or linearly uncorrelated variables.
 In other words, the axes of the dataset in question are reoriented so the Rayleigh quotient, which is related to the covariance of the dataset, along the first axis is maximized.
@@
 
-Methods:
+\section{Methods}
 
@@
 
-Results:
+\section{Results}
 

### 2016-05-11 scripts/plot_metrics.py
New data

@@ -4,69 +4,31 @@ import matplotlib.pyplot as plt
 import numpy as np
 import warnings
 import pandas
-from collections import defaultdict
 
-
-planner_strs = ["PRMstar", "RRTConnect"]
-pl_clrs = {"RRTstar": "r", "PRMstar": "b"}
-transformer_strs = ["FeatureAgglomeration", "TruncatedSVD",
-                    "PCA", "KernelPCA", "RandomizedPCA"]
-
-field_names = ["planner", "transformer", "n_collisions",
-               "planning_duration", "path_length",
-               "n_collisions_std", "planning_duration_std", "path_length_std",
-               "num_failed"]
-
-
-def load_data(filename):
-    data = np.genfromtxt(filename, delimiter=",", dtype=None)[1:]
-    planner_data = recursive_defaultdict(4, float)
-    for row in data:
-        for i in xrange(len(field_names) - 3):
-            planner_data[row[0]][
-                row[1]][row[-1]][field_names[i + 2]] = row[i + 2]
-    print planner_data
-    return planner_data
-
-
-def recursive_defaultdict(levels, dtype):
-    return recursive_defaultdict_helper(defaultdict(dtype), levels - 1)
-
-
-def recursive_defaultdict_helper(dd, level):
-    new_dd = defaultdict(lambda: dd)
-    if level == 1:
-        return new_dd
-    else:
-        return recursive_defaultdict_helper(new_dd, level - 1)
-
-
-def get_all(data, pl, n_ob, key):
-    xs = list()
-    for tr in transformer_strs:
-        xs.append(data[pl][tr][n_ob][key])
-    return xs
-
-
-def make_plots(data, key, n_ob):
-    rects = list()
-    fig, ax = plt.subplots()
-    ind = np.arange(len(transformer_strs))
-    width = 0.35
-    for i, pl in enumerate(planner_strs):
-        rect = ax.bar(
-            ind + i * width, get_all(data, pl, n_ob, key), width,
-            color=pl_clrs[pl],
-            label=pl)
-        rects.append(rect)
-        ax.set_ylabel(key)
-    ax.set_xticks(ind + width)
-    ax.set_xticklabels(transformer_strs)
-    plt.legend()
 
 if __name__ == "__main__":
     warnings.simplefilter("ignore")
[…]

### 2016-05-11 tests/experiments.py

@@ -16,28 +16,27 @@ from ompl import geometric as og
 from collections import defaultdict
 
 
-planners = [og.PRMstar]
+planners = [og.PRM, og.RRTConnect]
 transformers = [cluster.FeatureAgglomeration, decomp.TruncatedSVD,
                 decomp.PCA, decomp.RandomizedPCA,
                 None]
 
-planner_strs = ["PRMstar"]
+planner_strs = ["PRM", "RRTConnect"]
 transformer_strs = ["FeatureAgglomeration", "TruncatedSVD",
                     "PCA", "RandomizedPCA", "No Transform"]
 
 field_names = ["planner", "transformer", "n_collisions",
                "path_length", "n_collisions_std",
                "planning_duration_std", "path_length_std",
-               "num_failed", "n_obs", "is_full_dim"]
+               "num_failed", "n_obs", "is_full_dim", "duration",
+               "duration_std"]
 
-# n_obs = [70, 90, 110]
-n_obs = [14]
-n_runs = 2
-high_dim = 3
-low_dim = 3
+n_obs = [250, 300, 350]
+n_runs = 100
+high_dim = 8
+low_dim = 6
 timeout = 1.5
-rad_mean = 5
-# rad_mean = 7
+rad_mean = 14.8
 
 
 def path_size(path):
@@ -81,11 +80,11 @@ def run_experiments(filename):
                                     planner=pl, obstacles=obs,
                                     transform=tr)
                             try:
-                                path = planner.solve(start, goal, timeout)
-                                exit()
+                                path, dur = planner.solve(start, goal, timeout)
                                 data["n_collisions"].append(
                                     planner.check_path(path))
                                 data["path_length"].append(path_size(path))
+                                data["duration"].append(dur)
                             except ValueError:
                                 data["num_failed"].append(1)
                         except spatial.qhull.QhullError:
@@ -103,6 +102,8 @@ def run_experiments(filename):
                     row["path_length_std"] = np.std(data["path_length"])
                     row["num_failed"] = np.sum(data["num_failed"])
                     row["is_full_dim"] = tr is None
+                    row["duration"] = np.mean(data["duration"])
+                    row["duration_std"] = np.std(data["duration"])
                     writer.writerow(row)
 
 

### 2016-05-11 paper/paper.tex

@@
 state space for motion planning, which must take into account both the robot
-and obstacles, is called the configuration space \cite{lozano1983spatial}.
+and obstacles, is called the configuration space (or C-space) \cite{lozano1983spatial}.
 
@@
 The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
-PRM can be elevated to the asymptotically optimal PRM* by defining the radius to be a function of the number of vertices and the dimensionality of the space \cite{karaman2011sampling}.
+PRM can be elevated to the asymptotically optimal PRM*, which is described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of the number of vertices and the dimensionality of the space \cite{karaman2011sampling}.
 
-The Rapidly-expanding Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
-Starting at the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
-RRT* differs from RRT in that whenever a new node is added to the tree, connections are also tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
+The Rapidly exploring Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
+Starting at the initial configuration of the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
+RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
 The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.
@@
 
-We used Feature Agglomeration, Truncated SVD, PCA, Kernel PCA, Randomized PCA, and
+\section{Dimensionality Reduction}
+
+We used Feature Agglomeration, Truncated SVD, PCA, Randomized PCA, and
 our own ``Trained Johnson-Lindenstrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
 
-PCA: Principle Component Analysis, or PCA, transforms a set of variables into 
+\textbf{PCA}: Principle Component Analysis, or PCA, transforms a set of variables into 
 their ``principle components", or linearly uncorrelated variables.
@@
 
-Kernel PCA: Kernel PCA enables non-linear dimensionality reduction by using kernels to create a non-linear mapping from a high dimension to a lower dimension.
+%Kernel PCA: Kernel PCA enables non-linear dimensionality reduction by using kernels to create a non-linear mapping from a high dimension to a lower dimension.
 
-Randomized PCA: In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
+\textbf{Randomized PCA}: In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
 
-Feature Agglomeration: Feature Agglomeration groups together similar features to reduce the number of features.
+\textbf{Feature Agglomeration}: Feature Agglomeration groups together similar features to reduce the number of features.
 It does this by using agglomerative clustering, in which every data point starts as a cluster, and clusters are merged together.
 
-Truncated SVD: Truncated SVD works best with sparse matrices.
+\textbf{Truncated SVD}: Truncated SVD works best with sparse matrices.
 
-Trained Johnson-Lindenstrauss: Normal Johnson-Lindenstrauss uses a random mxn Gaussian matrix to reduce the dimensionality of a vector from n to m.
+\textbf{Trained Johnson-Lindenstrauss}: Normal Johnson-Lindenstrauss uses a random mxn Gaussian matrix to reduce the dimensionality of a vector from n to m.
 Instead of using a random matrix, we train a matrix on randomly-generated obstacles in n-space to minimize [MINIMIZE WHAT?]
@@
 Simplifications: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
-Therefore we limited ourselves to path planning for a point in $R^{n}$, and we approximate obstacles as being convex polytopes.
+Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximate obstacles as being convex polytopes.
 This is a valid simplification of the robot motion planning problem for several reasons.
-The first reason is that most real robots actually operate in a C-space similar to $R^{n}$.
-This is because $RP^{3}$, the space in which the rotation of a rigid body occurs, simplifies to $R^{3}$ if the joint cannot make a full rotation.
+The first reason is that most real robots actually operate in a C-space similar to $\mathbb{R}^n$.
+This is because $\mathbb{RP}^3$, the space in which the rotation of a rigid body occurs, simplifies to $\mathbb{R}^3$ if the joint cannot make a full rotation.
 Second, in defining the C-space, the obstacle space takes into account the interaction of the robot geometry with obstacles.
@@
 We implemented an n-dimensional world in which to test these algorithms.
-We structured our code so that we could run our tests in any space of dimension d >= 2.
-We also made visualization scripts for the 2- and 3-dimensional tests.
+We structured our code so that we could run our tests in any space of dimension $d \geq 2$.
+We also made visualization scripts for the 2- and 3-dimensional tests (examples of which can be seen in Figs ~\ref{fig:2d_examples} and ~\ref{fig:3d_examples}).
 
@@
 
-\begin{align}
+\begin{align*}
 x_1 &= r \cos(\phi_1) \\
@@
 x_n &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
-\end{align}
+\end{align*}
 
@@
     \end{subfigure}
+    \caption{2d examples}
+    \label{fig:2d_examples}
+\end{figure}
+
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{3d_example_1.pdf}
+        \caption{a}
+    \end{subfigure}%
+    ~ 
+    \begin{subfigure}[t]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{3d_example_2.pdf}
+        \caption{b}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{3d_example_3.pdf}
+        \caption{c}
+    \end{subfigure}
     \caption{3d examples}
@@
 \begin{algorithm}[ht] 
-    \caption{$\Function{Dimensionality Reduction}$}
-    \label{algo:DR}
-    \begin{algorithmic}[1]
-        \setcounter{ALC@line}{0}
-        \vspace*{1mm}
-
-        \STATE $Specify dimension, dimensionality reduction technique, and obstacle information$
-        \STATE $Generate obstacles based off of obstacle information$
-        \STATE $Fit dimensionality reduction transform to the space$
-        \STATE $Transform the space$
-        \STATE $Run PRM* over the space$
-        \STATE $Inverse transform the path back to the original dimension$
-        \STATE $Check for collisions$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[ht] 
     \caption{$\Function{PRM}^*$~\cite{karaman2011sampling}}
-    \label{algo:MS}
+    \label{alg:PRMstar}
     \begin{algorithmic}[1]
@@
     \caption{$\Function{RRT}^*$~\cite{karaman2011sampling}}
-    \label{algo:MS}
+    \label{alg:RRTstar}
     \begin{algorithmic}[1]
@@
 
+\begin{algorithm}[ht] 
+    \caption{$\Function{Dimensionality Reduction}$}
+    \label{alg:DR}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+
+        \STATE $Specify dimension, dimensionality reduction technique, and obstacle information$
+        \STATE $Generate obstacles based off of obstacle information$
+        \STATE $Fit dimensionality reduction transform to the space$
+        \STATE $Transform the space$
+        \STATE $Run PRM* over the space$
+        \STATE $Inverse transform the path back to the original dimension$
+        \STATE $Check for collisions$
+    \end{algorithmic}
+\end{algorithm}
+
 \bibliographystyle{plain}

### 2016-05-11 paper/paper.tex
Update paper with 2d figures
minor paper edit

@@
 
-Simplifications: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
+\textbe{Simplifications}: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
 Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximate obstacles as being convex polytopes.

### 2016-05-11 scripts/plot_metrics.py
Add plot to paper

@@ -1,7 +1,6 @@
 
 import seaborn as sns
 import matplotlib.pyplot as plt
-import numpy as np
 import warnings
 import pandas
 
@@ -10,23 +9,28 @@ if __name__ == "__main__":
     warnings.simplefilter("ignore")
     sns.set_context("poster", font_scale=1)
     data = pandas.read_csv("data/data.csv")
+    data.replace("FeatureAgglomeration", "F. Agglomeration", inplace=True)
     hue = "n_obs"
     rrt = data[data["planner"] == "RRTConnect"]
     prm = data[data["planner"] == "PRM"]
     plt.subplot(3, 2, 1)
     ax = sns.barplot(x="transformer", y="path_length", hue=hue, data=rrt)
     ax.set(xlabel="")
+    ax.legend().remove()
     plt.title("RRTConnect")
     plt.subplot(3, 2, 2)
     ax = sns.barplot(x="transformer", y="path_length", hue=hue, data=prm)
     ax.set(xlabel="")
+    ax.legend().remove()
     plt.title("PRM")
     plt.subplot(3, 2, 3)
     ax = sns.barplot(x="transformer", y="duration", hue=hue, data=rrt)
     ax.set(xlabel="")
+    ax.legend().remove()
     plt.subplot(3, 2, 4)
     ax = sns.barplot(x="transformer", y="duration", hue=hue, data=prm)
     ax.set(xlabel="")
+    ax.legend().remove()
     plt.subplot(3, 2, 5)
     sns.barplot(x="transformer", y="num_failed", hue=hue, data=rrt)
     plt.subplot(3, 2, 6)

### 2016-05-11 paper/paper.tex
Add plot to paper
paper mods

--- 2016-05-11 paper/paper.tex
@@
 
-Simplifications: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
+\textbf{Simplifications}: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
 Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximate obstacles as being convex polytopes.
@@
     \centering
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{3d_example_1.pdf}
+        \includegraphics[width=\textwidth]{2d_example_1.pdf}
         \caption{a}
@@
     ~ 
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{3d_example_2.pdf}
+        \includegraphics[width=\textwidth]{2d_example_2.pdf}
         \caption{b}
     \end{subfigure}
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[width=\textwidth]{3d_example_3.pdf}
+        \includegraphics[width=\textwidth]{2d_example_3.pdf}
         \caption{c}
@@
     \centering
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
@@
     ~ 
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
@@
     \end{subfigure}
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering

### 2016-05-13 paper/paper.tex
paper updates

@@
 the C-space are defined implicitly so that the path is constructed by randomly
-or pseudorandomly sampling points from the C-space and doing collision
-checking. In combinatorial motion planning, obstacles are defined explicitly
+or pseudorandomly sampling points from the C-space, doing collision
+checking, and connecting valid points to their nearest neighbors. In combinatorial motion planning, obstacles are defined explicitly
 and a complete search of the C-space is made. Because of the difficulty of
 explicitly defining all of the geometry in a potentially complex world,
-sampling-based motion planning techniques such as the Rapidly-exploring Random
-Tree (RRT) and Probabilistic RoadMap (PRM) planning are the most popular motion
+sampling-based motion planning techniques such as the Rapidly expanding Random
+Tree (RRT) and the Probabilistic RoadMap (PRM) are the most popular motion
 planning techniques.
 
-We used RRT* and PRM* in our project. The * indicates that these are the
-asymptotically optimal versions of RRT and PRM (they are guaranteed to find the
-optimal path as time goes to infinity).
+We used PRM* and RRTConnect in our project.
 
-The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations \cite{kavraki1996prm}.
+\textbf{PRM*}: The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations \cite{kavraki1996prm}.
 The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
 PRM can be elevated to the asymptotically optimal PRM*, which is described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of the number of vertices and the dimensionality of the space \cite{karaman2011sampling}.
+Since PRM* is asymptotically optimal, its solution will converge to the optimal path as time goes to infinity.
 
-The Rapidly exploring Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
-Starting at the initial configuration of the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
-RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
+\textbf{RRTConnect}: The Rapidly exploring Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
+Starting with the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
+The tree is ``rapidly exploring'' because the sample selection is biased towards areas with large Voronoi regions (where the Voronoi regions are computed based off of the nodes of the tree).
+RRTConnect, which is described in Alg~\ref{alg:RRTConnect} differs from RRT in that two trees, one with its root at the initial configuration and one with its root at the start configuration, grow towards each other \cite{kuffner2000rrt}.
+%RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
 The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{plots}
+    \caption{Figure}
+\end{figure}
+
 You can see examples of our code running in Fig ~\ref{fig:3d_examples}. The results are listed in Figure WHAT. They show that

### 2016-05-13 paper/paper.tex
added RRTConnect algorithm

@@
 \usepackage{mathabx}
+
+\usepackage{mathptmx}
+\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
 
@@
 
-\textbf{Simplifications}: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is extremely hard (specifying link geometry, defining constraints between links, etc).
-Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximate obstacles as being convex polytopes.
+\textbf{Simplifications}: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is very hard (specifying link geometry, defining constraints between links, etc).
+Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximated obstacles as being convex polytopes.
 This is a valid simplification of the robot motion planning problem for several reasons.
@@
 
-Next we made a random obstacle generator. We create n-dimensional convex polytopes by sampling $n^{2}$ random points on the n-sphere using the equations below.
+Next we made a random obstacle generator.
+We create n-dimensional convex polytopes by sampling $n^{2}$ random points on the n-sphere.
+We did this for each point by choosing $n-1$ random angles and then defining $x_1$ through $x_n$ using the equations below.
 
@@
 
-We then run Feature Agglomeration, Truncated SVD, PCA, Kernel PCA, and Randomized PCA to reduce the dimensionality of the space.
-We then run PRM* on each of these reduced dimensionality spaces, as well as on the full-dimensionality space as a control.
+We then ran Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to reduce the dimensionality of the space.
+We then ran PRM* and RRTConnect on each of these reduced dimensionality spaces, as well as on the full-dimensionality space as a control.
 We then run an inverse transform on the paths that we find to return them to the full-dimensionality space.
@@
 \begin{algorithm}[ht] 
-    \caption{$\Function{RRT}^*$~\cite{karaman2011sampling}}
-    \label{alg:RRTstar}
+    \caption{$\Function{EXTEND($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
+    \label{alg:EXTEND}
     \begin{algorithmic}[1]
@@
         \vspace*{1mm}
+        \STATE $ q_{near} \leftarrow \Function{NEAREST\_NEIGHBOR($q$,$\;\mathcal{T}$)} $
+        \IF{ $ \Function{NEW\_CONFIG($q$,$\;q_{near}$,$\;q_{new}$)} $}
+            \STATE $ \mathcal{T}.\text{add\_vertex($q_{new}$)}; $
+            \STATE $ \mathcal{T}.\text{add\_edge($q_{near}$,$\;q_{new}$)}; $
+            \IF{$ q_{new} = q $}
+                \RETURN $Reached$;
+            \ELSE
+                \RETURN $Advanced$;
+            \ENDIF
+        \ENDIF
+        \RETURN $Trapped$;
+    \end{algorithmic}
+\end{algorithm}
 
-        \STATE $V \leftarrow \{x_{\Var{init}}\}$
-        \STATE $E \leftarrow \emptyset$
-        \FOR{$i = 1 \ \TO \ n$}
-            \STATE $\xr \leftarrow \Var{SampleFree}_i$
-            \STATE $\xnearest \leftarrow \Function{Nearest}(G = (V, E), \xr)$
-            \STATE $\xnew \leftarrow \Function{Steer}(\xnearest, \xr)$
-            \IF{$\Function{ObstacleFree}(\xnearest, \xnew)$}
-                \STATE $\xnear \leftarrow \Function{Near}(G = (V, E), \xnew,
-                \min \{\gamma \cdot \Function{RRT}(\log(|V|) / |V|)^{1 / d}, \eta\})$
-                \STATE $V \leftarrow V \cup \{\xnew\}$
-                \STATE $\xmin \leftarrow \xnearest$
-                \STATE $\cmin \leftarrow \Function{Cost}(\xnear) +
-                c(\Function{Line}(\xnearest, \xnew))$
-                \FORALL{$\xnear \in X_{\Var{near}}$}
-                    \STATE $C \leftarrow \cost(\xnear) + \linef{\xnear, \xnew}$
-                    \IF{$\Function{CollisionFree}(\xnear, \xnew) \wedge C < \cmin$}
-                        \STATE $\xmin \leftarrow \xnear$
-                        \STATE $\cmin \leftarrow C$
+\begin{algorithm}[ht] 
+    \caption{$\Function{CONNECT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
+    \label{alg:CONNECT}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+        \REPEAT 
+            \STATE $S \leftarrow \Function{EXTEND($\mathcal{T}$,$\;q$)};$ 
+        \UNTIL {$ \textbf{not} \; (S = Advanced)$}
+        \RETURN $S;$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[ht] 
+    \caption{$\Function{RRTConnect($q_{init}$,$\;q_{goal}$)}$~\cite{kuffner2000rrt}}
+    \label{alg:RRTConnect}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+        \STATE $\mathcal{T}_a$.init($q_{init}$); $ \mathcal{T}_b$.init($q_{goal}$);
+        \FOR{$k = 1 \ \TO \ K$}
+            \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
+            \IF {$\textbf{not} \ (\Function{EXTEND($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
+                \IF {$ (\Function{CONNECT($\mathcal{T}_b,q_{new}$}) = Reached $}
+                    \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
                 \ENDIF
-                    \STATE $E \leftarrow E \cup \{(\xmin, \xnew)\}$
-                \ENDFOR
-                \FORALL{$\xnear \in X_{\Var{near}}$}
-                    \STATE $C \leftarrow \cost(\xnew) + \linef{\xnew, \xnear}$
-                    \IF{$\Function{CollisionFree}(\xnew, \xnear) \wedge C < \cost(\xnear)$}
-                        \STATE $\xparent \leftarrow \Function{Parent}(\xnear)$
-                        \STATE $E \leftarrow (E \backslash \{(\xparent, \xnear)\}) \cup \{(\xnew, \xnear)\}$
+            \STATE \Function{SWAP($\mathcal{T}_a,\mathcal{T}_b)$}
             \ENDIF
         \ENDFOR
-            \ENDIF
-        \ENDFOR
-        \RETURN $G = (V, E)$
-
+        \RETURN $Failure$
     \end{algorithmic}
 \end{algorithm}
+
+% \begin{algorithm}[ht] 
+%     \caption{$\Function{RRT}^*$~\cite{karaman2011sampling}}
+%     \label{alg:RRTConnect}
+%     \begin{algorithmic}[1]
+%         \setcounter{ALC@line}{0}
+%         \vspace*{1mm}
+
+%         \STATE $V \leftarrow \{x_{\Var{init}}\}$
+%         \STATE $E \leftarrow \emptyset$
+%         \FOR{$i = 1 \ \TO \ n$}
+%             \STATE $\xr \leftarrow \Var{SampleFree}_i$
+%             \STATE $\xnearest \leftarrow \Function{Nearest}(G = (V, E), \xr)$
+%             \STATE $\xnew \leftarrow \Function{Steer}(\xnearest, \xr)$
+%             \IF{$\Function{ObstacleFree}(\xnearest, \xnew)$}
+%                 \STATE $\xnear \leftarrow \Function{Near}(G = (V, E), \xnew,
+%                 \min \{\gamma \cdot \Function{RRT}(\log(|V|) / |V|)^{1 / d}, \eta\})$
+%                 \STATE $V \leftarrow V \cup \{\xnew\}$
+%                 \STATE $\xmin \leftarrow \xnearest$
+%                 \STATE $\cmin \leftarrow \Function{Cost}(\xnear) +
+%                 c(\Function{Line}(\xnearest, \xnew))$
+%                 \FORALL{$\xnear \in X_{\Var{near}}$}
+%                     \STATE $C \leftarrow \cost(\xnear) + \linef{\xnear, \xnew}$
+%                     \IF{$\Function{CollisionFree}(\xnear, \xnew) \wedge C < \cmin$}
+%                         \STATE $\xmin \leftarrow \xnear$
+%                         \STATE $\cmin \leftarrow C$
+%                     \ENDIF
+%                     \STATE $E \leftarrow E \cup \{(\xmin, \xnew)\}$
+%                 \ENDFOR
+%                 \FORALL{$\xnear \in X_{\Var{near}}$}
+%                     \STATE $C \leftarrow \cost(\xnew) + \linef{\xnew, \xnear}$
+%                     \IF{$\Function{CollisionFree}(\xnew, \xnear) \wedge C < \cost(\xnear)$}
+%                         \STATE $\xparent \leftarrow \Function{Parent}(\xnear)$
+%                         \STATE $E \leftarrow (E \backslash \{(\xparent, \xnear)\}) \cup \{(\xnew, \xnear)\}$
+%                     \ENDIF
+%                 \ENDFOR
+%             \ENDIF
+%         \ENDFOR
+%         \RETURN $G = (V, E)$
+
+%     \end{algorithmic}
+% \end{algorithm}
 

### 2016-05-13 paper/paper.tex
I will update the results section tmrw

@@
 The tree is ``rapidly exploring'' because the sample selection is biased towards areas with large Voronoi regions (where the Voronoi regions are computed based off of the nodes of the tree).
-RRTConnect, which is described in Alg~\ref{alg:RRTConnect} differs from RRT in that two trees, one with its root at the initial configuration and one with its root at the start configuration, grow towards each other \cite{kuffner2000rrt}.
+RRTConnect, which is described in Algs~\ref{alg:extend_RRT},~\ref{alg:connect_RRT}, and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root at the initial configuration and one with its root at the start configuration, grow towards each other \cite{kuffner2000rrt}.
 %RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
@@
 
-We then ran Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to reduce the dimensionality of the space.
+After generating random obstacles, we ran Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to reduce the dimensionality of the space.
 We then ran PRM* and RRTConnect on each of these reduced dimensionality spaces, as well as on the full-dimensionality space as a control.
-We then run an inverse transform on the paths that we find to return them to the full-dimensionality space.
-Then we do collision checking and check the number of collisions the path makes with the obstacles.
+We then ran an inverse transform on the paths that we find to return them to the full-dimensionality space.
+We collected data on the number of runs that fail, the time it takes to run the algorithm, and the path length returned.
+A summary of this paragraph can be found in Alg~\ref{alg:DR}.
 
@@
 \begin{algorithm}[ht] 
-    \caption{$\Function{EXTEND($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
-    \label{alg:EXTEND}
+    \caption{$\Function{Extend\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
+    \label{alg:extend_RRT}
     \begin{algorithmic}[1]
@@
 \begin{algorithm}[ht] 
-    \caption{$\Function{CONNECT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
-    \label{alg:CONNECT}
+    \caption{$\Function{Connect\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
+    \label{alg:connect_RRT}
     \begin{algorithmic}[1]
@@
         \REPEAT 
-            \STATE $S \leftarrow \Function{EXTEND($\mathcal{T}$,$\;q$)};$ 
+            \STATE $S \leftarrow \Function{EXTEND\_RRT($\mathcal{T}$,$\;q$)};$ 
         \UNTIL {$ \textbf{not} \; (S = Advanced)$}
@@
             \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
-            \IF {$\textbf{not} \ (\Function{EXTEND($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
-                \IF {$ (\Function{CONNECT($\mathcal{T}_b,q_{new}$}) = Reached $}
+            \IF {$\textbf{not} \ (\Function{EXTEND\_RRT($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
+                \IF {$ (\Function{CONNECT\_RRT($\mathcal{T}_b,q_{new}$}) = Reached $}
                     \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
@@
 
-        \STATE $Specify dimension, dimensionality reduction technique, and obstacle information$
-        \STATE $Generate obstacles based off of obstacle information$
-        \STATE $Fit dimensionality reduction transform to the space$
-        \STATE $Transform the space$
-        \STATE $Run PRM* over the space$
-        \STATE $Inverse transform the path back to the original dimension$
-        \STATE $Check for collisions$
+        \STATE Specify dimension, dimensionality reduction technique, and obstacle information
+        \STATE Generate random obstacles based off of obstacle information
+        \STATE Fit dimensionality reduction transform to the space
+        \STATE Transform the space
+        \STATE Run PRM* or RRTConnect over the space
+        \STATE Inverse transform the path back to the original dimension
+        \STATE Check for collisions
+        \RETURN Path\_length, duration, num\_failed
     \end{algorithmic}

### 2016-05-13 flatland/__init__.py
Paper stuff maybe sorta kinda no

@@ -1,10 +1,12 @@
 
 __all__ = ["FLSolution", "FLPlanner", "RandomObstacleGen",
-           "make_path_plot_2d", "make_path_plot_3d", "DRPlanner"]
+           "make_path_plot_2d", "make_path_plot_3d", "DRPlanner",
+           "TrainedJL"]
 
 from planner import FLSolution
 from planner import FLPlanner
 from dr import DRPlanner
+from trainedjl import TrainedJL
 from obstacles import RandomObstacleGen
 from plot_path import make_path_plot_2d
 from plot_path import make_path_plot_3d

### 2016-05-13 flatland/dr.py

@@ -4,7 +4,6 @@ import planner
 import numpy as np
 import polytope
 import sklearn.cluster as cluster
-import sklearn.decomposition as decomp
 import scipy.spatial as spatial
 from ompl import geometric as og
 
@@ -71,36 +70,43 @@ class DRPlanner(object):
                     collision_count += 1
         return collision_count
 
-    def solve(self, st, gl, timeout=1.0):
+    def solve_low_dim(self, st, gl, timeout):
+        tr = self.fit(st, gl)
+        exit()
+        ld_obsts = self.low_dim_obstacles(tr)
+        ld_st_gl = tr.transform(np.array([st, gl]))
+        flpl = planner.FLPlanner(
+            dim=self.low_dim,
+            planner=self.pl,
+            obstacles=ld_obsts,
+            low_bound=self.low_bound,
+            high_bound=self.high_bound)
+        start = time.time()
+        res = flpl.solve(ld_st_gl[0], ld_st_gl[1], timeout)
+        end = time.time()
+        ld_path = res.get_solution()
+        hd_path = tr.inverse_transform(self.path_to_arr(ld_path))
+        hd_path = np.vstack((st, hd_path, gl))
+        return hd_path, end - start
+
+    def solve_high_dim(self, st, gl, timeout=1.0):
+        flpl = planner.FLPlanner(
+            dim=self.high_dim,
+            planner=self.pl,
+            obstacles=self.obstacles,
+            low_bound=self.low_bound,
+            high_bound=self.high_bound)
+        start = time.time()
+        res = flpl.solve(st, gl, timeout)
+        end = time.time()
+        res.write_to_file("sandbox/prmpath.txt")
+        flpl.save_obstacles()
+        path = res.get_solution()
+        route = self.path_to_arr(path)
+        return route, end - start
+
+    def solve(self, st, gl, timeout):
         if self.low_dim < self.high_dim:
-            tr = self.fit(st, gl)
-            ld_obsts = self.low_dim_obstacles(tr)
-            ld_st_gl = tr.transform(np.array([st, gl]))
-            flpl = planner.FLPlanner(
-                dim=self.low_dim,
-                planner=self.pl,
-                obstacles=ld_obsts,
-                low_bound=self.low_bound,
-                high_bound=self.high_bound)
-            start = time.time()
-            res = flpl.solve(ld_st_gl[0], ld_st_gl[1], timeout)
-            end = time.time()
[…]

### 2016-05-13 flatland/trainedjl.py

@@ -1,0 +1,37 @@
+
+import numpy as np
+import scipy.optimize as opt
+import random
+
+
+class TrainedJL(object):
+
+    def __init__(self, n_components=2):
+        self.low_dim = n_components
+
+    def fit(self, X):
+        self.high_dim = X.shape[1]
+        self.X = X
+        x0 = np.ones((1, self.high_dim * self.low_dim))
+        cons = ({"type": "ineq", "fun": lambda x: np.all(x != 0)})
+        res = opt.minimize(self.objective, x0, constraints=cons,
+                           method="SLSQP")
+        print res
+        return res.x
+
+    def objective(self, x):
+        err = 0
+        mat = x.reshape((self.high_dim, self.low_dim))
+        # # i = random.randint(0, self.X.shape[0])
+        # # j = random.randint(0, self.X.shape[0])
+        # i = 1
+        # j = 1000
+        # # print i, j
+        for i in xrange(self.X.shape[0] / 2000):
+            for j in xrange(self.X.shape[0] / 2000):
+                lxi = np.dot(self.X[i], mat)
+                lxj = np.dot(self.X[j], mat)
+                h_dist = np.linalg.norm(self.X[i] - self.X[j])
+                l_dist = np.linalg.norm(lxi - lxj)
+                err += abs(h_dist - l_dist)
+        return err

### 2016-05-13 tests/experiments.py

@@ -16,10 +16,13 @@ from ompl import geometric as og
 from collections import defaultdict
 
 
-planners = [og.PRM, og.RRTConnect]
-transformers = [cluster.FeatureAgglomeration, decomp.TruncatedSVD,
-                decomp.PCA, decomp.RandomizedPCA,
-                None]
+# planners = [og.PRM, og.RRTConnect]
+planners = [og.PRM]
+# transformers = [cluster.FeatureAgglomeration, decomp.TruncatedSVD,
+#                 decomp.PCA, decomp.RandomizedPCA,
+#                 None]
+transformers = [flatland.TrainedJL]
+
 
 planner_strs = ["PRM", "RRTConnect"]
 transformer_strs = ["FeatureAgglomeration", "TruncatedSVD",

### 2016-05-13 paper/paper.tex
Adding some changes to the report

@@
 
-\textbf{PRM*}: The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations \cite{kavraki1996prm}.
+\subsection{PRM*}: The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations \cite{kavraki1996prm}.
 The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
@@
 
-\textbf{RRTConnect}: The Rapidly exploring Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
+\subsection{RRTConnect}: The Rapidly exploring Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
 Starting with the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
@@
 
-We used Feature Agglomeration, Truncated SVD, PCA, Randomized PCA, and
-our own ``Trained Johnson-Lindenstrauss" to perform dimensionality reduction. We will briefly summarize each of these techniques.
+We used Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to perform dimensionality reduction. We will briefly summarize each of these techniques.
 
-\textbf{PCA}: Principle Component Analysis, or PCA, transforms a set of variables into 
+\subsection{PCA} Principle Component Analysis, or PCA, transforms a set of variables into 
 their ``principle components", or linearly uncorrelated variables.
@@
 
-\textbf{Randomized PCA}: In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
+\subsection{Randomized PCA} In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
 
-\textbf{Feature Agglomeration}: Feature Agglomeration groups together similar features to reduce the number of features.
+\subsection{Feature Agglomeration} Feature Agglomeration groups together similar features to reduce the number of features.
 It does this by using agglomerative clustering, in which every data point starts as a cluster, and clusters are merged together.
 
-\textbf{Truncated SVD}: Truncated SVD works best with sparse matrices.
-
-\textbf{Trained Johnson-Lindenstrauss}: Normal Johnson-Lindenstrauss uses a random mxn Gaussian matrix to reduce the dimensionality of a vector from n to m.
-Instead of using a random matrix, we train a matrix on randomly-generated obstacles in n-space to minimize [MINIMIZE WHAT?]
+\subsection{Truncated SVD} Truncated SVD works best with sparse matrices.
 
@@
 
-\textbf{Simplifications}: Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is very hard (specifying link geometry, defining constraints between links, etc).
+\subsection{Simplifications} Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is very hard (specifying link geometry, defining constraints between links, etc).
 Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximated obstacles as being convex polytopes.

### 2016-05-13 paper/paper.tex
Update paper

@@
 \usepackage[pdftex]{graphicx}
-\graphicspath{ {figs/} }
+\graphicspath{{figs/}}
 \usepackage{subcaption}
@@
 described with the 3-dimensional Special Euclidean group (known as SE(3)).
-SE(3) is homeomorphic to the topological space $\mathbb{R}^3\bigtimes\mathbb{RP}^3$ (where $\mathbb{RP}^3$ is the
+SE(3) is homeomorphic to the topological space
+$\mathbb{R}^3\bigtimes\mathbb{RP}^3$ (where $\mathbb{RP}^3$ is the
 3-dimensional real projective plane). Therefore it is easy to imagine a
-multi-link robot \cite{kuindersma2015optimization} or a multi-robot system \cite{alonso2015multi} with an arbitrarily complex state space. The picture is
+multi-link robot \cite{kuindersma2015optimization} or a multi-robot system
+\cite{alonso2015multi} with an arbitrarily complex state space. The picture is
 further complicated by the addition of obstacles into a robot's world. The
 state space for motion planning, which must take into account both the robot
-and obstacles, is called the configuration space (or C-space) \cite{lozano1983spatial}.
+and obstacles, is called the configuration space (or C-space)
+\cite{lozano1983spatial}.
 
@@
 which can be divided into two classes, sampling-based motion planning and
-combinatorial motion planning \cite{lavalle2006planning}. In sampling-based motion planning, obstacles in
-the C-space are defined implicitly so that the path is constructed by randomly
-or pseudorandomly sampling points from the C-space, doing collision
-checking, and connecting valid points to their nearest neighbors. In combinatorial motion planning, obstacles are defined explicitly
+combinatorial motion planning \cite{lavalle2006planning}. In sampling-based
+motion planning, obstacles in the C-space are defined implicitly so that the
+path is constructed by randomly or pseudorandomly sampling points from the
+C-space, doing collision checking, and connecting valid points to their nearest
+neighbors. In combinatorial motion planning, obstacles are defined explicitly
 and a complete search of the C-space is made. Because of the difficulty of
@@
 
-\subsection{PRM*}: The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in which a roadmap is constructed by sampling random points in the C-space, and the query phase in which a graph search over the roadmap is made to connect initial and goal configurations \cite{kavraki1996prm}.
-The map is constructed during the preprocessing phase by sampling random points in the C-space and then adding them to the vertices in the roadmap within a certain radius of the point.
-PRM can be elevated to the asymptotically optimal PRM*, which is described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of the number of vertices and the dimensionality of the space \cite{karaman2011sampling}.
-Since PRM* is asymptotically optimal, its solution will converge to the optimal path as time goes to infinity.
+\subsection{PRM*}
 
-\subsection{RRTConnect}: The Rapidly exploring Random Tree algorithm works by building up a tree from the initial configuration to the goal configuration \cite{lavalle1998rrt}.
-Starting with the initial configuration as the root, it incrementally attempts to add randomly selected samples to the tree until it reaches the goal configuration.
-The tree is ``rapidly exploring'' because the sample selection is biased towards areas with large Voronoi regions (where the Voronoi regions are computed based off of the nodes of the tree).
-RRTConnect, which is described in Algs~\ref{alg:extend_RRT},~\ref{alg:connect_RRT}, and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root at the initial configuration and one with its root at the start configuration, grow towards each other \cite{kuffner2000rrt}.
+The Probabilistic RoadMap algorithm has two phases: the preprocessing phase in
+which a roadmap is constructed by sampling random points in the C-space, and
+the query phase in which a graph search over the roadmap is made to connect
+initial and goal configurations \cite{kavraki1996prm}. The map is constructed
+during the preprocessing phase by sampling random points in the C-space and
+then adding them to the vertices in the roadmap within a certain radius of the
+point. PRM can be elevated to the asymptotically optimal PRM*, which is
+described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of
+the number of vertices and the dimensionality of the space
+\cite{karaman2011sampling}. Since PRM* is asymptotically optimal, its solution
+will converge to the optimal path as time goes to infinity.
+
+\subsection{RRTConnect}
+
+The Rapidly exploring Random Tree algorithm works by building up a tree from
+the initial configuration to the goal configuration \cite{lavalle1998rrt}.
+Starting with the initial configuration as the root, it incrementally attempts
+to add randomly selected samples to the tree until it reaches the goal
+configuration.  The tree is ``rapidly exploring'' because the sample selection
+is biased towards areas with large Voronoi regions (where the Voronoi regions
+are computed based off of the nodes of the tree). RRTConnect, which is
+described in Algs~\ref{alg:extend_RRT},~\ref{alg:connect_RRT},
+and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root
+at the initial configuration and one with its root at the start configuration,
+grow towards each other \cite{kuffner2000rrt}.
 %RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
@@
 
-We used Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to perform dimensionality reduction. We will briefly summarize each of these techniques.
+We used Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to
+perform dimensionality reduction. We will briefly summarize each of these
+techniques.
 
-\subsection{PCA} Principle Component Analysis, or PCA, transforms a set of variables into 
-their ``principle components", or linearly uncorrelated variables.
-In other words, the axes of the dataset in question are reoriented so the Rayleigh quotient, which is related to the covariance of the dataset, along the first axis is maximized.
-After the first axis has been set, the second axis is then set to maximize the Rayleigh quotient of the dataset along that axis, and so on, so that the first axis has the highest Rayleigh quotient, the second axis has the second highest, and so on.
-Dimensionality reduction is then performed by chopping off the axes with the lowest Rayleigh quotients (which corresponded to axes with the lowest variance).
+\subsection{PCA}
+
+Principle Component Analysis, or PCA, transforms a set of variables into their
+``principle components", or linearly uncorrelated variables. In other words,
+the axes of the dataset in question are reoriented so the Rayleigh quotient,
+which is related to the covariance of the dataset, along the first axis is
+maximized. After the first axis has been set, the second axis is then set to
+maximize the Rayleigh quotient of the dataset along that axis, and so on, so
+that the first axis has the highest Rayleigh quotient, the second axis has the
+second highest, and so on. Dimensionality reduction is then performed by
+chopping off the axes with the lowest Rayleigh quotients (which corresponded to
+axes with the lowest variance).
 
@@
 
-%Kernel PCA: Kernel PCA enables non-linear dimensionality reduction by using kernels to create a non-linear mapping from a high dimension to a lower dimension.
+\subsection{Randomized PCA}
 
-\subsection{Randomized PCA} In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
+In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
 
-\subsection{Feature Agglomeration} Feature Agglomeration groups together similar features to reduce the number of features.
-It does this by using agglomerative clustering, in which every data point starts as a cluster, and clusters are merged together.
+\subsection{Feature Agglomeration}
 
-\subsection{Truncated SVD} Truncated SVD works best with sparse matrices.
+Feature Agglomeration groups together similar features to reduce the number of
+features. It does this by using agglomerative clustering, in which every data
+point starts as a cluster, and clusters are merged together.
+
+\subsection{Truncated SVD}
+
+Truncated SVD works best with sparse matrices.
 
@@
 
-\subsection{Simplifications} Although our aim is to show how dimensionality reduction techniques can speed up motion planning for robots, doing motion planning for an actual robot is very hard (specifying link geometry, defining constraints between links, etc).
-Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$, and we approximated obstacles as being convex polytopes.
-This is a valid simplification of the robot motion planning problem for several reasons.
-The first reason is that most real robots actually operate in a C-space similar to $\mathbb{R}^n$.
-This is because $\mathbb{RP}^3$, the space in which the rotation of a rigid body occurs, simplifies to $\mathbb{R}^3$ if the joint cannot make a full rotation.
-Second, in defining the C-space, the obstacle space takes into account the interaction of the robot geometry with obstacles.
-Therefore motion planning through the free space can be done using a point to represent the robot (since the geometry of the robot is implicitly included in the obstacle space).
-Therefore, although our simulations do not represent real robots and real obstacles, they do approximate them.
+\subsection{Simplifications}
 
-We implemented an n-dimensional world in which to test these algorithms.
-We structured our code so that we could run our tests in any space of dimension $d \geq 2$.
-We also made visualization scripts for the 2- and 3-dimensional tests (examples of which can be seen in Figs ~\ref{fig:2d_examples} and ~\ref{fig:3d_examples}).
+Although our aim is to show how dimensionality reduction techniques can speed
+up motion planning for robots, doing motion planning for an actual robot is
+very hard (specifying link geometry, defining constraints between links, etc).
+Therefore we limited ourselves to path planning for a point in $\mathbb{R}^n$,
+and we approximated obstacles as being convex polytopes. This is a valid
+simplification of the robot motion planning problem for several reasons. The
+first reason is that most real robots actually operate in a C-space similar to
+$\mathbb{R}^n$. This is because $\mathbb{RP}^3$, the space in which the
+rotation of a rigid body occurs, simplifies to $\mathbb{R}^3$ if the joint
+cannot make a full rotation. Second, in defining the C-space, the obstacle
+space takes into account the interaction of the robot geometry with obstacles.
+Therefore motion planning through the free space can be done using a point to
+represent the robot (since the geometry of the robot is implicitly included in
+the obstacle space). Therefore, although our simulations do not represent real
+robots and real obstacles, they do approximate them.
 
-Next we made a random obstacle generator.
-We create n-dimensional convex polytopes by sampling $n^{2}$ random points on the n-sphere.
-We did this for each point by choosing $n-1$ random angles and then defining $x_1$ through $x_n$ using the equations below.
+We implemented an n-dimensional world in which to test these algorithms. We
+structured our code so that we could run our tests in any space of dimension $d
+\geq 2$. We also made visualization scripts for the 2- and 3-dimensional tests
+(examples of which can be seen in Figs~\ref{fig:2d_examples}
+and~\ref{fig:3d_examples}).
+
+Next we made a random obstacle generator. We create n-dimensional convex
+polytopes by sampling $n^{2}$ random points on the n-sphere. We did this for
+each point by choosing $n-1$ random angles and then defining $x_1$ through
+$x_n$ using the equations below.
 
@@
 
-After generating random obstacles, we ran Feature Agglomeration, Truncated SVD, PCA, and Randomized PCA to reduce the dimensionality of the space.
-We then ran PRM* and RRTConnect on each of these reduced dimensionality spaces, as well as on the full-dimensionality space as a control.
-We then ran an inverse transform on the paths that we find to return them to the full-dimensionality space.
-We collected data on the number of runs that fail, the time it takes to run the algorithm, and the path length returned.
-A summary of this paragraph can be found in Alg~\ref{alg:DR}.
+After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
+PCA, and Randomized PCA to reduce the dimensionality of the space. We then ran
+PRM* and RRTConnect on each of these reduced dimensionality spaces, as well as
+on the full-dimensionality space as a control. We then ran an inverse transform
+on the paths that we find to return them to the full-dimensionality space. We
+collected data on the number of runs that fail, the time it takes to run the
+algorithm, and the path length returned. A summary of this paragraph can be
+found in Alg~\ref{alg:DR}.
 
@@
     \end{subfigure}%
-    ~ 
     \begin{subfigure}[t]{0.32\textwidth}
@@
 
-You can see examples of our code running in Fig ~\ref{fig:3d_examples}. The results are listed in Figure WHAT. They show that
+You can see examples of our code running in Fig~\ref{fig:3d_examples}. The
+results are listed in Figure WHAT. They show that
 

### 2016-05-13 paper/paper.tex
made plots huge

@@
 
+\subsection{Experimental Setup}
+
 We implemented an n-dimensional world in which to test these algorithms. We
 structured our code so that we could run our tests in any space of dimension $d
-\geq 2$. We also made visualization scripts for the 2- and 3-dimensional tests
-(examples of which can be seen in Figs~\ref{fig:2d_examples}
-and~\ref{fig:3d_examples}).
+\geq 2$. We also made visualization scripts for the 2 and 3 dimensional tests.
+Examples of randomly generated 2D and 3D scenes with candidate solution paths
+can be seen in Fig.~\ref{fig:examples}.
 
@@
 
-\begin{figure}[t!]
+\begin{figure}[h!]
     \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{2d_example_1.pdf}
-        \caption{a}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{2d_example_2.pdf}
-        \caption{b}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{2d_example_3.pdf}
-        \caption{c}
-    \end{subfigure}
-    \caption{2d examples}
-    \label{fig:2d_examples}
-\end{figure}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{3d_example_1.pdf}
-        \caption{a}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{3d_example_2.pdf}
-        \caption{b}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{3d_example_3.pdf}
-        \caption{c}
-    \end{subfigure}
-    \caption{3d examples}
-    \label{fig:3d_examples}
+    \includegraphics[width=0.32\textwidth]{2d_example_1.pdf}
+    \includegraphics[width=0.32\textwidth]{2d_example_2.pdf}
+    \includegraphics[width=0.32\textwidth]{2d_example_3.pdf} \\
+    \includegraphics[width=0.32\textwidth]{3d_example_1.pdf}
+    \includegraphics[width=0.32\textwidth]{3d_example_2.pdf}
+    \includegraphics[width=0.32\textwidth]{3d_example_3.pdf}
+    \caption{Examples of paths found in 2D and 3D scenes of random obstacles
+    using the PRM algorithm. The green and red circles indicate the initial
+    and goal configurations respectively. The blue path indicates the candidate
+    solution through the space that connects the initial and goal
+    configurations.}
+    \label{fig:examples}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\textwidth]{plots}
+    \makebox[\textwidth][c]{\includegraphics[width=1.5\textwidth]{plots}}
     \caption{Figure}
@@
 
-You can see examples of our code running in Fig~\ref{fig:3d_examples}. The
+You can see examples of our code running in Fig.~\ref{fig:examples}. The
 results are listed in Figure WHAT. They show that

### 2016-05-13 paper/paper.tex
updates to paper

--- 2016-05-13 paper/paper.tex
@@
 Principle Component Analysis, or PCA, transforms a set of variables into their
-``principle components", or linearly uncorrelated variables. In other words,
-the axes of the dataset in question are reoriented so the Rayleigh quotient,
-which is related to the covariance of the dataset, along the first axis is
-maximized. After the first axis has been set, the second axis is then set to
-maximize the Rayleigh quotient of the dataset along that axis, and so on, so
-that the first axis has the highest Rayleigh quotient, the second axis has the
-second highest, and so on. Dimensionality reduction is then performed by
-chopping off the axes with the lowest Rayleigh quotients (which corresponded to
-axes with the lowest variance).
-
-$$
-\mathbf{w}_{(1)} = {\operatorname{\arg\,max}}\, \left\{ \frac{\mathbf{w}^T\mathbf{X}^T \mathbf{X w}}{\mathbf{w}^T \mathbf{w}} \right\}
-$$
+``principal components", or linearly uncorrelated variables \cite{shlens2014pca}.
+Given an $m \times n$ matrix $\mathbf{X}$, you can find its principal components by finding an orthonormal matrix $\mathbf{P}$ in $ \mathbf{Y} = \mathbf{PX} $ such that $\mathbf{C_Y} = \mathbf{YY}^T/n$.
+The rows of $\mathbf{P}$ are the principal components of $\mathbf{X}$.
+In other words,
+the axes of the dataset in question are reoriented so that the covariance of the data along the first axis is minimized. After the first axis has been set, the second axis is then set to
+minimize the covariance of the dataset along that axis, and so on, so
+that the first axis has the lowest covariance, the second axis has the
+second lowest, and so on. Dimensionality reduction is then performed by
+chopping off the axes with the high covariance.
 
@@
 
-In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
+In Randomized PCA, computations are sped up by approximating the largest singular vectors of the components.
 
@@
 
+Feature Agglomeration is a form of hierarchical clustering in which clusters are merged or split successively \cite{featureagglomeration}.
 Feature Agglomeration groups together similar features to reduce the number of
@@
 
-Truncated SVD works best with sparse matrices.
+In singular value decomposition, a matrix $\mathbf{M}$ is factorized into the form $\mathbf{U} \Sigma \mathbf{V^*}$.
+$\Sigma$ is a diagonal matrix, and the entries of $\Sigma$ are the singular values of $\mathbf{M}$.
+In Truncated SVD, $\mathbf{M}$ is approximated by matrices $\mathbf{U_k} \Sigma_k \mathbf{V^*_k}$, where $ k < r $, $r$ being the number of non-zero singular values of the matrix \cite{maciejewski1989svd}.
+In effect you are approximating $\mathbf{M}$ by cutting off low-value singular values.
+In this sense Truncated SVD is similar to PCA; however, instead of cutting off the lowest eigenvalues of the coviarance matrix, you are cutting off the singular values of the matrix.
 

### 2016-05-13 paper/paper.tex
Merge conflicts to follow

--- 2016-05-13 paper/paper.tex
@@
 are computed based off of the nodes of the tree). RRTConnect, which is
-described in Algs~\ref{alg:extend_RRT},~\ref{alg:connect_RRT},
+described in Algos.~\ref{alg:extend_RRT},~\ref{alg:connect_RRT},
 and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root
@@
 
-In Randomized PCA, computations are sped up by only approximating the singular vectors of the components.
+In Randomized PCA, computations are sped up by only approximating the singular
+vectors of the components.
 
@@
 Next we made a random obstacle generator. We create n-dimensional convex
-polytopes by sampling $n^{2}$ random points on the n-sphere. We did this for
-each point by choosing $n-1$ random angles and then defining $x_1$ through
-$x_n$ using the equations below.
+polytopes by sampling $n^{2}$ random points within the $n$-sphere. To generate
+a single point, you must sample $n - 1$ angles from a uniform distribution
+between 0 and $2\pi$, $n$ radii from a Gaussian distribution,
+$\mathcal{N}(r_{\mu}, r_{\sigma})$ where $r_{\mu}$ and $r_{\sigma}$ are the
+mean and standard deviation for the radius of the convex polytope, and $n$
+floats to act as the center of the $n$-sphere.  Then you use the series of
+equations in Eq.~\ref{eq:nsphere} to produce a randomly sampled point within
+the $n$-sphere where $c_k$, $r_k$, and $\phi_k$ are the centers, radii, and
+angles respectively.  To produce a convex polytope, we then computed the
+$n$-dimensional convex hull over these points. The convex hull then provided
+the normal equations and offsets for each facet on the polytope.
 
-\begin{align*}
-x_1 &= r \cos(\phi_1) \\
-x_2 &= r \sin(\phi_1) \cos(\phi_2) \\
-x_3 &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \\
+\begin{align}
+    \begin{split}
+        x_1 &= c_1 + r_1 \cos(\phi_1) \\
+        x_2 &= c_2 + r_2 \sin(\phi_1) \cos(\phi_2) \\
+        x_3 &= c_3 + r_3 \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \\
             &\vdots\\
-x_{n-1} &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \cos(\phi_{n-1}) \\
-x_n &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
-\end{align*}
+        x_{n-1} &= c_{n - 1} + r_{n - 1} \sin(\phi_1) \cdots \sin(\phi_{n-2}) \cos(\phi_{n-1}) \\
+        x_n &= c_n + r_{n} \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
+    \end{split}
+    \label{eq:nsphere}
+\end{align}
 
 After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
-PCA, and Randomized PCA to reduce the dimensionality of the space. We then ran
-PRM* and RRTConnect on each of these reduced dimensionality spaces, as well as
-on the full-dimensionality space as a control. We then ran an inverse transform
-on the paths that we find to return them to the full-dimensionality space. We
-collected data on the number of runs that fail, the time it takes to run the
-algorithm, and the path length returned. A summary of this paragraph can be
-found in Alg~\ref{alg:DR}.
+PCA, and Randomized PCA to reduce the dimensionality of the space by using the
+sampled points within each $n$-sphere as training data. We then ran PRM* and
+RRTConnect on each of these reduced dimensionality spaces, as well as on the
+fully-dimensional space as a control. We then ran an inverse transform on the
+paths that we find to return them to the fully-dimensional space in order to
+verify that the path in the reduced dimension did not come into collision with
+obstacles. We collected data on the number of runs that fail to find a
+solution, the time it takes to run the algorithm, and the length of the
+solution if one was found.  A summary of this paragraph can be found in
+Algo.~\ref{alg:DR}.
+
+For our experiments, we varied the number of random obstacles in the scene, the
+planning algorithm used, and the method of dimensionality reduction. For each
+problem instance, the random generator was seeded in order to make fair
+comparisons between the planners and dimensionality reduction techniques.
+Also, since the algorithms are non-deterministic, each problem instance
+executed 100 times. We used scenes with 250, 300, and 350 obstacles in order to
+measure how the planners and dimensionality reduction techniques vary in their
+performance with varying levels of obstacle density. In the results presented,
+we are reducing from 8 dimensions to 6 dimensions and using mean radius of
+$14.8$ for the convex polytopes.
 

### 2016-05-13 paper/paper.tex

@@
 Principle Component Analysis, or PCA, transforms a set of variables into their
-``principle components", or linearly uncorrelated variables. In other words,
-the axes of the dataset in question are reoriented so the Rayleigh quotient,
-which is related to the covariance of the dataset, along the first axis is
-maximized. After the first axis has been set, the second axis is then set to
-maximize the Rayleigh quotient of the dataset along that axis, and so on, so
-that the first axis has the highest Rayleigh quotient, the second axis has the
-second highest, and so on. Dimensionality reduction is then performed by
-chopping off the axes with the lowest Rayleigh quotients (which corresponded to
-axes with the lowest variance).
-
-$$
-\mathbf{w}_{(1)} = {\operatorname{\arg\,max}}\, \left\{ \frac{\mathbf{w}^T\mathbf{X}^T \mathbf{X w}}{\mathbf{w}^T \mathbf{w}} \right\}
-$$
+``principal components", or linearly uncorrelated variables
+\cite{shlens2014pca}. Given an $m \times n$ matrix $\mathbf{X}$, you can find
+its principal components by finding an orthonormal matrix $\mathbf{P}$ in $
+\mathbf{Y} = \mathbf{PX} $ such that $\mathbf{C_Y} = \mathbf{YY}^T/n$. The rows
+of $\mathbf{P}$ are the principal components of $\mathbf{X}$. In other words,
+the axes of the dataset in question are reoriented so that the covariance of
+the data along the first axis is minimized. After the first axis has been set,
+the second axis is then set to minimize the covariance of the dataset along
+that axis, and so on, so that the first axis has the lowest covariance, the
+second axis has the second lowest, and so on. Dimensionality reduction is then
+performed by chopping off the axes with the high covariance.
 
@@
 
-Feature Agglomeration groups together similar features to reduce the number of
+Feature Agglomeration is a form of hierarchical clustering in which clusters
+are merged or split successively \cite{featureagglomeration}. Feature
+Agglomeration groups together similar features to reduce the number of
 features. It does this by using agglomerative clustering, in which every data
@@
 
-Truncated SVD works best with sparse matrices.
+In singular value decomposition, a matrix $\mathbf{M}$ is factorized into the
+form $\mathbf{U} \Sigma \mathbf{V^*}$. $\Sigma$ is a diagonal matrix, and the
+entries of $\Sigma$ are the singular values of $\mathbf{M}$. In Truncated SVD,
+$\mathbf{M}$ is approximated by matrices $\mathbf{U_k} \Sigma_k
+\mathbf{V^*_k}$, where $ k < r $, $r$ being the number of non-zero singular
+values of the matrix \cite{maciejewski1989svd}. In effect you are approximating
+$\mathbf{M}$ by cutting off low-value singular values. In this sense Truncated
+SVD is similar to PCA; however, instead of cutting off the lowest eigenvalues
+of the coviarance matrix, you are cutting off the singular values of the
+matrix.
 

### 2016-05-13 paper/paper.tex
got rid of prm*

--- 2016-05-13 paper/paper.tex
@@
 
-We used PRM* and RRTConnect in our project.
+We used PRM and RRTConnect in our project.
 
-\subsection{PRM*}
+\subsection{PRM}
 
@@
 then adding them to the vertices in the roadmap within a certain radius of the
-point. PRM can be elevated to the asymptotically optimal PRM*, which is
-described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of
-the number of vertices and the dimensionality of the space
-\cite{karaman2011sampling}. Since PRM* is asymptotically optimal, its solution
-will converge to the optimal path as time goes to infinity.
+point. The algorithm is described in Alg~\ref{alg:PRM}.
+%PRM can be elevated to the asymptotically optimal PRM*, which is
+%described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of
+%the number of vertices and the dimensionality of the space
+%\cite{karaman2011sampling}. Since PRM* is asymptotically optimal, its solution
+%will converge to the optimal path as time goes to infinity.
 
@@
 \begin{algorithm}[ht] 
-    \caption{$\Function{PRM}^*$~\cite{karaman2011sampling}}
-    \label{alg:PRMstar}
+    \caption{$\Function{PRM}~\cite{karaman2011sampling}}
+    \label{alg:PRM}
     \begin{algorithmic}[1]
@@
         \FORALL{$v \in V$}
-            \STATE $U \leftarrow \Function{Near}(G = (V, E), v,
-            \gamma \cdot \Function{PRM}(\log(n) / n) ^ {1 / d})
-            \backslash \{v\}$
-            \FORALL{$u \in U$}
+            \STATE $U \leftarrow \Function{Near}(G = (V, E), x_{rand}, r);$
+            \STATE $ V \leftarrow V \cup {x_{rand}} $
+            \FORALL{$u \in U $}
                 \IF{$\Function{CollisionFree}(v, u)$}

### 2016-05-13 paper/paper.tex
Fixed prm
rearranged algorithms
writing upr esults

@@
 
-\section{Background}
+\section{Introduction}
 
@@
 point. The algorithm is described in Alg~\ref{alg:PRM}.
-%PRM can be elevated to the asymptotically optimal PRM*, which is
-%described in Alg~\ref{alg:PRMstar}, by defining the radius to be a function of
-%the number of vertices and the dimensionality of the space
-%\cite{karaman2011sampling}. Since PRM* is asymptotically optimal, its solution
-%will converge to the optimal path as time goes to infinity.
 
@@
 grow towards each other \cite{kuffner2000rrt}.
-%RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}.
-The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.
 
@@
 matrix.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.32\textwidth]{2d_example_1.pdf}
+    \includegraphics[width=0.32\textwidth]{2d_example_2.pdf}
+    \includegraphics[width=0.32\textwidth]{2d_example_3.pdf} \\
+    \includegraphics[width=0.32\textwidth]{3d_example_1.pdf}
+    \includegraphics[width=0.32\textwidth]{3d_example_2.pdf}
+    \includegraphics[width=0.32\textwidth]{3d_example_3.pdf}
+    \caption{Examples of paths found in 2D and 3D scenes of random obstacles
+    using the PRM algorithm. The green and red circles indicate the initial
+    and goal configurations respectively. The blue path indicates the candidate
+    solution through the space that connects the initial and goal
+    configurations.}
+    \label{fig:examples}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=0.32\textwidth]{2d_example_1.pdf}
-    \includegraphics[width=0.32\textwidth]{2d_example_2.pdf}
-    \includegraphics[width=0.32\textwidth]{2d_example_3.pdf} \\
-    \includegraphics[width=0.32\textwidth]{3d_example_1.pdf}
-    \includegraphics[width=0.32\textwidth]{3d_example_2.pdf}
-    \includegraphics[width=0.32\textwidth]{3d_example_3.pdf}
-    \caption{Examples of paths found in 2D and 3D scenes of random obstacles
-    using the PRM algorithm. The green and red circles indicate the initial
-    and goal configurations respectively. The blue path indicates the candidate
-    solution through the space that connects the initial and goal
-    configurations.}
-    \label{fig:examples}
+    \makebox[\textwidth][c]{\includegraphics[width=1.5\textwidth]{plots}}
+
+    \caption{Plots showing the length of the solution paths if found, the
+    time needed to find the solution, and the number of failed attempts out of
+    100 for each method of dimensionality reduction as well as in the high
+    dimensional space. The column of plots on the left shows the results for
+    the RRTConnect algorithm and the plots on the right show the results for
+PRM. Note, The Y-axes limits are not equal and that abbreviations were used to
+save space.  F.  Agg.  stands for Feature Agglomeration and R. PCA stands for
+Randomized PCA.}
+
+    \label{fig:plots}
+
 \end{figure}
 
-\begin{figure}[h!]
-    \centering
-    \makebox[\textwidth][c]{\includegraphics[width=1.5\textwidth]{plots}}
-    \caption{Figure}
-\end{figure}
+Fig.~\ref{fig:plots} shows plots of various measures of performance using
+different methods of dimensionality reduction for motion planing. The left and
+right columns show the results for RRTConnect and PRM algorithms respectively.
+The first, second, and third row show the average path length, computational
+time in seconds, and number of failed runs out of 100 respectively. For each
+dimensionality reduction technique, there are three bars presenting the results
+for 250, 300, and 350 randomly generated obstacles respectively. From the data
+we can see that both planners produced similarly sized paths for each
+dimensionality reduction technique, and that by not reducing the dimension, the
+planner is able to find a path roughly two times shorter.
 
-You can see examples of our code running in Fig.~\ref{fig:examples}. The
-results are listed in Figure WHAT. They show that
+In terms of computational time, there is no significant difference between
+Truncated SVD, PCA, Randomized PCA, planning in the original space when
+planning with the RRTConnect algorithm. Whereas using Feature Agglomeration
+took slightly longer to compute for each set number of obstacles. When using
+PRM, the computational time was significantly improved by reducing the
+dimension for all methods.  Based on the data, using PRM without dimensionality
+reduction took roughly twice as long to compute for each method. This differing
+result between PRM and RRTConnect is likely due to the fact that PRM is a
+multi-query planning algorithm meaning that it constructs a data-structure that
+can be used for multiple planning queries whereas RRTConnect is only looking to
+find a single path. This means that the PRM algorithm tries to fully represent
+the connectivity of the environment whilst RRTConnect needs to determine the
+connectivity directed towards a goal. The reason dimensionality reduction
+significantly improves the computational time for PRM is that it is much faster
+to exhaust the search space at lower dimensions as opposed to higher
+dimensions.
+
+The RRTConnect algorithm had a much lower failure rate than PRM in all
+respective dimensionality reduction technique. This is due to the fact that
+RRTConnect is growing a tree towards the goal and PRM is randomly sampling and
+connecting points to a growing graph. For both algorithms, a similar trend
+exists for failure rate for planning in the reduced dimension space.
 
 \begin{algorithm}[ht] 
-    \caption{$\Function{PRM}~\cite{karaman2011sampling}}
+    \caption{$\Function{PRM}$~\cite{karaman2011sampling}}
     \label{alg:PRM}
@@
         \FORALL{$v \in V$}
-            \STATE $U \leftarrow \Function{Near}(G = (V, E), x_{rand}, r);$
-            \STATE $ V \leftarrow V \cup {x_{rand}} $
+            \STATE $U \leftarrow \Function{NearestNeighbors}(v)$
             \FORALL{$u \in U $}

### 2016-05-13 scripts/plot_metrics.py

@@ -3,7 +3,6 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import warnings
 import pandas
-import sys
 
 
 def prettify(text):
@@ -20,6 +19,21 @@ if __name__ == "__main__":
     hue = "n_obs"
     rrt = data[data["planner"] == "RRTConnect"]
     prm = data[data["planner"] == "PRM"]
+    n_m = prm.query("transformer == 'No Transform'")["duration"].mean()
+    print n_m / prm.query("transformer == 'F. Agg.'")["duration"].mean()
+    print n_m / prm.query("transformer == 'TruncatedSVD'")["duration"].mean()
+    print n_m / prm.query("transformer == 'PCA'")["duration"].mean()
+    print n_m / prm.query("transformer == 'R. PCA'")["duration"].mean()
+
+    print "----"
+
+    n_m = prm.query("transformer == 'No Transform'")["path_length"].mean()
+    print prm.query("transformer == 'F. Agg.'")["path_length"].mean() / n_m
+    print prm.query("transformer == 'TruncatedSVD'")["path_length"].mean() \
+        / n_m
+    print prm.query("transformer == 'PCA'")["path_length"].mean() / n_m
+    print prm.query("transformer == 'R. PCA'")["path_length"].mean() / n_m
+
     plt.subplot(3, 2, 1)
     ax = sns.barplot(x="transformer", y="path_length", hue=hue, data=rrt)
     ax.set(xlabel="")
@@ -34,7 +48,7 @@ if __name__ == "__main__":
     plt.title("PRM")
     plt.subplot(3, 2, 3)
     ax = sns.barplot(x="transformer", y="duration", hue=hue, data=rrt)
-    ax.set(ylabel="Duration")
+    ax.set(ylabel="Comp. Time [s]")
     ax.set(xlabel="")
     ax.legend().remove()
     plt.subplot(3, 2, 4)

### 2016-05-14 paper/paper.tex
Finished results section

@@
 
+\begin{algorithm}[h!] 
+    \caption{$\Function{PRM}$~\cite{karaman2011sampling}}
+    \label{alg:PRM}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+
+        \STATE $V \leftarrow \{x_{\Var{init}}\} \cup \{\Var{SampleFree}_i\}_{i=1,\ldots,n}$
+        \STATE $E \leftarrow \emptyset$
+        \FORALL{$v \in V$}
+            \STATE $U \leftarrow \Function{NearestNeighbors}(v)$
+            \FORALL{$u \in U $}
+                \IF{$\Function{CollisionFree}(v, u)$}
+                    \STATE $E \leftarrow E \cup \{(v, u), (u, v)\}$
+                \ENDIF
+            \ENDFOR
+        \ENDFOR
+        \RETURN $G = (V, E)$
+
+    \end{algorithmic}
+\end{algorithm}
+
 \subsection{RRTConnect}
+
+\begin{algorithm}[h!] 
+    \caption{$\Function{Extend\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
+    \label{alg:extend_RRT}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+        \STATE $ q_{near} \leftarrow \Function{NEAREST\_NEIGHBOR($q$,$\;\mathcal{T}$)} $
+        \IF{ $ \Function{NEW\_CONFIG($q$,$\;q_{near}$,$\;q_{new}$)} $}
+            \STATE $ \mathcal{T}.\text{add\_vertex($q_{new}$)}; $
+            \STATE $ \mathcal{T}.\text{add\_edge($q_{near}$,$\;q_{new}$)}; $
+            \IF{$ q_{new} = q $}
+                \RETURN $Reached$;
+            \ELSE
+                \RETURN $Advanced$;
+            \ENDIF
+        \ENDIF
+        \RETURN $Trapped$;
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[h!] 
+    \caption{$\Function{Connect\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
+    \label{alg:connect_RRT}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+        \REPEAT 
+            \STATE $S \leftarrow \Function{EXTEND\_RRT($\mathcal{T}$,$\;q$)};$ 
+        \UNTIL {$ \textbf{not} \; (S = Advanced)$}
+        \RETURN $S;$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[!h] 
+    \caption{$\Function{RRTConnect($q_{init}$,$\;q_{goal}$)}$~\cite{kuffner2000rrt}}
+    \label{alg:RRTConnect}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+        \STATE $\mathcal{T}_a$.init($q_{init}$); $ \mathcal{T}_b$.init($q_{goal}$);
+        \FOR{$k = 1 \ \TO \ K$}
+            \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
+            \IF {$\textbf{not} \ (\Function{EXTEND\_RRT($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
+                \IF {$ (\Function{CONNECT\_RRT($\mathcal{T}_b,q_{new}$}) = Reached $}
+                    \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
+                \ENDIF
+            \STATE \Function{SWAP($\mathcal{T}_a,\mathcal{T}_b)$}
+            \ENDIF
+        \ENDFOR
+        \RETURN $Failure$
+    \end{algorithmic}
+\end{algorithm}
 
@@
 
+\begin{algorithm}[h!] 
+    \caption{$\Function{Dimensionality Reduction}$}
+    \label{alg:DR}
+    \begin{algorithmic}[1]
+        \setcounter{ALC@line}{0}
+        \vspace*{1mm}
+
+        \STATE Specify dimension, dimensionality reduction technique, and obstacle information
+        \STATE Generate random obstacles based off of obstacle information
+        \STATE Fit dimensionality reduction transform to the space
+        \STATE Transform the space
+        \STATE Run PRM* or RRTConnect over the space
+        \STATE Inverse transform the path back to the original dimension
+        \STATE Check for collisions
+        \RETURN Path\_length, duration, num\_failed
+    \end{algorithmic}
+\end{algorithm}
+
 After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
@@
 solution, the time it takes to run the algorithm, and the length of the
-solution if one was found.  A summary of this paragraph can be found in
+solution if one was found. Note, that the length of the path is computed after
+it has been transformed back into the original high dimensional space in order
+to make a fair comparison. A summary of this paragraph can be found in
 Algo.~\ref{alg:DR}.
@@
 
-\section{Results}
+\section{Results \& Discussion}
+
+Fig.~\ref{fig:plots} shows plots of various measures of performance using
+different methods of dimensionality reduction for motion planing. Please note
+that the Y-axes of these plots are not equivalent in the left and right column.
+The left and right columns show the results for RRTConnect and PRM algorithms
+respectively.  The first, second, and third row show the average path length,
+computational time in seconds, and number of failed runs out of 100
+respectively. For each dimensionality reduction technique, there are three bars
+presenting the results for 250, 300, and 350 randomly generated obstacles
+respectively. From the data we can see that both planners produced similarly
+sized paths for each dimensionality reduction technique, and that by not
+reducing the dimension, the planner is able to find a path roughly two times
+shorter.
 
@@
 \end{figure}
-
-Fig.~\ref{fig:plots} shows plots of various measures of performance using
-different methods of dimensionality reduction for motion planing. The left and
-right columns show the results for RRTConnect and PRM algorithms respectively.
-The first, second, and third row show the average path length, computational
-time in seconds, and number of failed runs out of 100 respectively. For each
-dimensionality reduction technique, there are three bars presenting the results
-for 250, 300, and 350 randomly generated obstacles respectively. From the data
-we can see that both planners produced similarly sized paths for each
-dimensionality reduction technique, and that by not reducing the dimension, the
-planner is able to find a path roughly two times shorter.
 
@@
 connecting points to a growing graph. For both algorithms, a similar trend
-exists for failure rate for planning in the reduced dimension space.
+exists for failure rates when planning in the reduced dimension space. For both
+planners, Feature Agglomeration failed more frequently than the rest of the
+techniques, and planning in the high dimensional space had the smallest failure
+rate. The higher failure rate for Feature Agglomeration is because Feature
+Agglomeration works best when features from the data can be combined due to
+repetition and redundancy. Since this data was in the $\mathbb{R}^n$ and the
+random obstacles were generated all over the space, Feature Agglomeration
+failed to produce an embedding that also worked for planning in the lower
+dimensional space. Aside from Feature Agglomeration, the dimensionality
+reduction techniques failed to find a path to the goal configuration between 20
+and 60 percent of the time for PRM and between 5 and 35 percent for RRTConnect.
+This is due to the projection of the higher dimensional obstacles into a lower
+dimension causing occlusions in the lower dimensional space where they would
+not exist in the higher dimension. However, these dimensionality reduction
+techniques can be seen as independent random variables for each run, therefore,
+you would be able to run the algorithm again with the same probability of
+failure. As the number of obstacles increased, so did the failure rate because
+the lower dimensional spaces become more densely populated with obstacles
+causing more occlusions.
 
-\begin{algorithm}[ht] 
-    \caption{$\Function{PRM}$~\cite{karaman2011sampling}}
-    \label{alg:PRM}
-    \begin{algorithmic}[1]
-        \setcounter{ALC@line}{0}
-        \vspace*{1mm}
+From the data, we can see that dimensionality reduction techniques can reduce
+the optimality of the planning algorithm, but in multi-query planning
+algorithms, these techniques can significantly reduce the time needed to
+compute a solution. Based on the data, it seems that the Truncated SVD
+algorithm can be used to produce the lowest number of failures and produces
+similar results in the path length and computation time  as compared to the
+other techniques used. Also, we Can see that Feature Agglomeration fails to
+find a solution path more frequently than the other techniques, but produces
+shorter paths when it does find a path.
 
-        \STATE $V \leftarrow \{x_{\Var{init}}\} \cup \{\Var{SampleFree}_i\}_{i=1,\ldots,n}$
-        \STATE $E \leftarrow \emptyset$
-        \FORALL{$v \in V$}
-            \STATE $U \leftarrow \Function{NearestNeighbors}(v)$
-            \FORALL{$u \in U $}
-                \IF{$\Function{CollisionFree}(v, u)$}
-                    \STATE $E \leftarrow E \cup \{(v, u), (u, v)\}$
-                \ENDIF
-            \ENDFOR
-        \ENDFOR
-        \RETURN $G = (V, E)$
-
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[ht] 
-    \caption{$\Function{Extend\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
-    \label{alg:extend_RRT}
-    \begin{algorithmic}[1]
-        \setcounter{ALC@line}{0}
-        \vspace*{1mm}
-        \STATE $ q_{near} \leftarrow \Function{NEAREST\_NEIGHBOR($q$,$\;\mathcal{T}$)} $
-        \IF{ $ \Function{NEW\_CONFIG($q$,$\;q_{near}$,$\;q_{new}$)} $}
-            \STATE $ \mathcal{T}.\text{add\_vertex($q_{new}$)}; $
-            \STATE $ \mathcal{T}.\text{add\_edge($q_{near}$,$\;q_{new}$)}; $
-            \IF{$ q_{new} = q $}
-                \RETURN $Reached$;
-            \ELSE
-                \RETURN $Advanced$;
-            \ENDIF
-        \ENDIF
-        \RETURN $Trapped$;
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[ht] 
-    \caption{$\Function{Connect\_RRT($\mathcal{T}$,$q$)}$~\cite{kuffner2000rrt}}
-    \label{alg:connect_RRT}
-    \begin{algorithmic}[1]
-        \setcounter{ALC@line}{0}
-        \vspace*{1mm}
-        \REPEAT 
-            \STATE $S \leftarrow \Function{EXTEND\_RRT($\mathcal{T}$,$\;q$)};$ 
-        \UNTIL {$ \textbf{not} \; (S = Advanced)$}
-        \RETURN $S;$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[ht] 
-    \caption{$\Function{RRTConnect($q_{init}$,$\;q_{goal}$)}$~\cite{kuffner2000rrt}}
-    \label{alg:RRTConnect}
-    \begin{algorithmic}[1]
-        \setcounter{ALC@line}{0}
-        \vspace*{1mm}
-        \STATE $\mathcal{T}_a$.init($q_{init}$); $ \mathcal{T}_b$.init($q_{goal}$);
-        \FOR{$k = 1 \ \TO \ K$}
-            \STATE $q_{rand} \leftarrow \Function{RANDOM\_CONFIG()}$;
-            \IF {$\textbf{not} \ (\Function{EXTEND\_RRT($\mathcal{T}_a$,$\;q_{rand} = Trapped$)})$}
-                \IF {$ (\Function{CONNECT\_RRT($\mathcal{T}_b,q_{new}$}) = Reached $}
-                    \RETURN $ \Function{PATH($\mathcal{T}_a,\mathcal{T}_b$)} $
-                \ENDIF
-            \STATE \Function{SWAP($\mathcal{T}_a,\mathcal{T}_b)$}
-            \ENDIF
-        \ENDFOR
-        \RETURN $Failure$
-    \end{algorithmic}
-\end{algorithm}
-
-% \begin{algorithm}[ht] 
-%     \caption{$\Function{RRT}^*$~\cite{karaman2011sampling}}
-%     \label{alg:RRTConnect}
-%     \begin{algorithmic}[1]
-%         \setcounter{ALC@line}{0}
-%         \vspace*{1mm}
-
-%         \STATE $V \leftarrow \{x_{\Var{init}}\}$
-%         \STATE $E \leftarrow \emptyset$
-%         \FOR{$i = 1 \ \TO \ n$}
-%             \STATE $\xr \leftarrow \Var{SampleFree}_i$
-%             \STATE $\xnearest \leftarrow \Function{Nearest}(G = (V, E), \xr)$
-%             \STATE $\xnew \leftarrow \Function{Steer}(\xnearest, \xr)$
-%             \IF{$\Function{ObstacleFree}(\xnearest, \xnew)$}
-%                 \STATE $\xnear \leftarrow \Function{Near}(G = (V, E), \xnew,
-%                 \min \{\gamma \cdot \Function{RRT}(\log(|V|) / |V|)^{1 / d}, \eta\})$
-%                 \STATE $V \leftarrow V \cup \{\xnew\}$
-%                 \STATE $\xmin \leftarrow \xnearest$
-%                 \STATE $\cmin \leftarrow \Function{Cost}(\xnear) +
-%                 c(\Function{Line}(\xnearest, \xnew))$
-%                 \FORALL{$\xnear \in X_{\Var{near}}$}
-%                     \STATE $C \leftarrow \cost(\xnear) + \linef{\xnear, \xnew}$
-%                     \IF{$\Function{CollisionFree}(\xnear, \xnew) \wedge C < \cmin$}
-%                         \STATE $\xmin \leftarrow \xnear$
-%                         \STATE $\cmin \leftarrow C$
-%                     \ENDIF
-%                     \STATE $E \leftarrow E \cup \{(\xmin, \xnew)\}$
-%                 \ENDFOR
-%                 \FORALL{$\xnear \in X_{\Var{near}}$}
-%                     \STATE $C \leftarrow \cost(\xnew) + \linef{\xnew, \xnear}$
-%                     \IF{$\Function{CollisionFree}(\xnew, \xnear) \wedge C < \cost(\xnear)$}
-%                         \STATE $\xparent \leftarrow \Function{Parent}(\xnear)$
-%                         \STATE $E \leftarrow (E \backslash \{(\xparent, \xnear)\}) \cup \{(\xnew, \xnear)\}$
-%                     \ENDIF
-%                 \ENDFOR
-%             \ENDIF
-%         \ENDFOR
-%         \RETURN $G = (V, E)$
-
-%     \end{algorithmic}
-% \end{algorithm}
-
-\begin{algorithm}[ht] 
-    \caption{$\Function{Dimensionality Reduction}$}
-    \label{alg:DR}
-    \begin{algorithmic}[1]
-        \setcounter{ALC@line}{0}
-        \vspace*{1mm}
-
-        \STATE Specify dimension, dimensionality reduction technique, and obstacle information
-        \STATE Generate random obstacles based off of obstacle information
-        \STATE Fit dimensionality reduction transform to the space
-        \STATE Transform the space
-        \STATE Run PRM* or RRTConnect over the space
-        \STATE Inverse transform the path back to the original dimension
-        \STATE Check for collisions
-        \RETURN Path\_length, duration, num\_failed
-    \end{algorithmic}
-\end{algorithm}
+\section{Conclusion}
 

### 2016-05-14 paper/paper.tex
finished conclusion

@@
         \STATE Transform the space
-        \STATE Run PRM* or RRTConnect over the space
+        \STATE Run PRM or RRTConnect over the space
         \STATE Inverse transform the path back to the original dimension
@@
 PCA, and Randomized PCA to reduce the dimensionality of the space by using the
-sampled points within each $n$-sphere as training data. We then ran PRM* and
+sampled points within each $n$-sphere as training data. We then ran PRM and
 RRTConnect on each of these reduced dimensionality spaces, as well as on the
 fully-dimensional space as a control. We then ran an inverse transform on the
-paths that we find to return them to the fully-dimensional space in order to
+paths that we found to return them to the fully-dimensional space in order to
 verify that the path in the reduced dimension did not come into collision with
@@
 solution, the time it takes to run the algorithm, and the length of the
-solution if one was found. Note, that the length of the path is computed after
+solution if one was found. Note that the length of the path is computed after
 it has been transformed back into the original high dimensional space in order
@@
 that the Y-axes of these plots are not equivalent in the left and right column.
-The left and right columns show the results for RRTConnect and PRM algorithms
+The left and right columns show the results for the RRTConnect and PRM algorithms
 respectively.  The first, second, and third row show the average path length,
@@
 In terms of computational time, there is no significant difference between
-Truncated SVD, PCA, Randomized PCA, planning in the original space when
-planning with the RRTConnect algorithm. Whereas using Feature Agglomeration
+Truncated SVD, PCA, Randomized PCA, and planning in the original space when
+planning with the RRTConnect algorithm, whereas using Feature Agglomeration
 took slightly longer to compute for each set number of obstacles. When using
@@
 result between PRM and RRTConnect is likely due to the fact that PRM is a
-multi-query planning algorithm meaning that it constructs a data-structure that
-can be used for multiple planning queries whereas RRTConnect is only looking to
+multi-query planning algorithm, meaning that it constructs a data-structure that
+can be used for multiple planning queries, whereas RRTConnect is only looking to
 find a single path. This means that the PRM algorithm tries to fully represent
@@
 Agglomeration works best when features from the data can be combined due to
-repetition and redundancy. Since this data was in the $\mathbb{R}^n$ and the
+repetition and redundancy. Since this data was in $\mathbb{R}^n$ and the
 random obstacles were generated all over the space, Feature Agglomeration
@@
 not exist in the higher dimension. However, these dimensionality reduction
-techniques can be seen as independent random variables for each run, therefore,
+techniques can be seen as independent random variables for each run; therefore,
 you would be able to run the algorithm again with the same probability of
 failure. As the number of obstacles increased, so did the failure rate because
-the lower dimensional spaces become more densely populated with obstacles
+the lower dimensional spaces become more densely populated with obstacles,
 causing more occlusions.
@@
 similar results in the path length and computation time  as compared to the
-other techniques used. Also, we Can see that Feature Agglomeration fails to
+other techniques used. Also, we can see that Feature Agglomeration fails to
 find a solution path more frequently than the other techniques, but produces
@@
 
+Thus, in this project we compared several dimensionality reduction techniques against each other and against the option of no reduction.
+Our experiments showed that, at least for the case of random obstacles in $\mathbb{R}^n$, Feature Agglomeration performs very poorly, while Truncated SVD slightly outperforms PCA and Randomized PCA.
+We also found that, for multi-query planning algorithms, dimensionality reduction significantly reduced the calculation time.
+However, this benefit was not seen for the single-query RRTConnect.
+Unfortunately, dimensionality reduction seems to come at a cost - in our tests, dimensionality reduction doubled path length and often caused the planning algorithm to fail.
+In conclusion, in our tests, dimensionality reduction improved the computation time of the multi-query planning algorithm but resulted in longer path lengths and more failures than the nonreduced-dimension case.
+
 \bibliographystyle{plain}
